### 2024-12-25 index.qmd
samkc

# Preface {.unnumbered}

This is a Quarto book.

scs

To learn more about Quarto books visit <https://quarto.org/docs/books>.

```{r}
1 + 1
```