### 2025-03-24 appendix1.qmd
Appendices with individual numbering added

---
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
---

# Supplementary materials 1

Some supplementary materials.