### 2025-03-24 article1.qmd
Yalm structure and base qmd created. To populate with the actual articles.

---
title: "Creating frames of reference for chert exploitation during the Late Pleistocene in Southwesternmost Iberia"
---