### 2025-03-24 article2.qmd

---
title: "Within and Beyond: Chert Procurement Patterns During The Upper Palaeolithic in Southwesternmost Iberia"
---