### 2025-03-24 index.qmd

# Acknowledgments {.unnumbered}

This is a Quarto book.

To learn more about Quarto books visit <https://quarto.org/docs/books>.

```{r}
1 + 1
```