### 2025-03-24 thesis2024.tex
First render

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  a4paper,
  DIV=11,
  numbers=noendperiod]{scrreprt}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{colortbl}
\usepackage{array}
\usepackage{anyfontsize}
\usepackage{multirow}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{bookmark}{}{\usepackage{bookmark}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={thesis2024},
  pdfauthor={Joana Belmiro},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{thesis2024}
\author{Joana Belmiro}
\date{2024-12-25}

\begin{document}
\maketitle

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}

\setstretch{2}
\bookmarksetup{startatroot}

\chapter*{Acknowledgments}\label{acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

\markboth{Acknowledgments}{Acknowledgments}

This is a Quarto book.

To learn more about Quarto books visit
\url{https://quarto.org/docs/books}.

\begin{Shaded}
\begin{Highlighting}[]
\DecValTok{1} \SpecialCharTok{+} \DecValTok{1}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 2
\end{verbatim}

\bookmarksetup{startatroot}

\chapter{Abstract}\label{abstract}

In summary, this book has no content whatsoever.

\begin{Shaded}
\begin{Highlighting}[]
\DecValTok{1} \SpecialCharTok{+} \DecValTok{1}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 2
\end{verbatim}

\bookmarksetup{startatroot}

\chapter{Introduction}\label{introduction}

This is a book created from markdown and executable code.

See Knuth (1984) for additional discussion of literate programming.

\begin{Shaded}
\begin{Highlighting}[]
\DecValTok{1} \SpecialCharTok{+} \DecValTok{1}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 2
\end{verbatim}

\bookmarksetup{startatroot}

\chapter{Creating frames of reference for chert exploitation during the
Late Pleistocene in Southwesternmost
Iberia}\label{creating-frames-of-reference-for-chert-exploitation-during-the-late-pleistocene-in-southwesternmost-iberia}

\bookmarksetup{startatroot}

\chapter{Within and beyond: chert procurement patterns during the Upper
Palaeolithic in Southwesternmost
Iberia}\label{within-and-beyond-chert-procurement-patterns-during-the-upper-palaeolithic-in-southwesternmost-iberia}

\bookmarksetup{startatroot}

\chapter{Within and beyond: chert procurement patterns during the Upper
Palaeolithic in Southwesternmost
Iberia}\label{within-and-beyond-chert-procurement-patterns-during-the-upper-palaeolithic-in-southwesternmost-iberia-1}

\textbf{Authors}: Joana Belmiro, Xavier Terradas, Salvador
Dominguez-Bella, João Cascalheira

\newpage

\section*{Abstract}\label{abstract-1}
\addcontentsline{toc}{section}{Abstract}

\markright{Abstract}

Analyses of raw materials and the distinction between local/regional and
long-distance sources have proven invaluable for understanding the
extensive movements, interactions, and social networks during the Upper
Palaeolithic in the Iberian Peninsula. However, unlike other parts of
Iberia, research on the management and acquisition of raw materials in
the south and west of Iberia remains relatively underdeveloped. Despite
significant knowledge about the technological practices of Palaeolithic
hunter-gatherers from southern Portugal, particularly from studies
conducted at the site of Vale Boi, there is a noticeable lack of focus
on raw materials management. This paper presents the first comprehensive
characterization of chert raw materials from the Gravettian,
Proto-Solutrean, and Solutrean occupations at Vale Boi, using both
macroscopic and petrographic techniques. Our study reveals that the
majority of chert found at Vale Boi originates locally, within a 20 km
radius. However, a non-negligible portion of the chert comes from
non-local sources, indicating \textgreater200 km raw material
circulation from central Portugal and southern Spain.

\textbf{keywords:} Upper Paleolithic; Iberian Peninsula; Lithic raw
materials; Petrography

\newpage

\section{Introduction}\label{introduction-1}

Knappable raw materials play a crucial role in understanding the
mobility and lifeways of past hunter-gatherers, given the ubiquitous
presence of lithic artefacts throughout Prehistory. More than mere rocks
with which stone tools were produced, the management of these resources
is intimately connected to a group's technological, social, and cultural
organisation, potentially influencing the group's overall survivability
(Oestmo 2017; Binford 1979; Bleed 1986; Bousman 1993; Torrence 1983;
Gould and Saggers 1985).

Several key topics in the study of hunter-gatherer behaviour and
organisation have been explored in different geographic regions and
chronological periods through raw material analysis, such as modalities
of procurement and mobility strategies (Ambrose and Lorenz 1990; Binford
1979; Binford and Stone 1985; Gould 1985; Gould and Saggers 1985; Kuhn
1991; McCall 2007), occupation types and their duration (Kuhn 2004;
Surovell 2009), the establishment and dimension of social networks
(Whallon 2006), as well as exchanges between groups or individuals
(Gamble 1999). These analyses are often achieved by systematically
characterizing geological and archaeological raw materials and
establishing correlations between samples from both origins. The
ultimate aim is to identify the corresponding sources, enabling, for
example, a precise differentiation between local and non-local raw
materials, and examine their distribution within a site over different
occupations and periods.

These approaches and concepts have also been extensively and
successfully applied to the study of lithic assemblages from prehistoric
archaeological sites in the Iberian Peninsula (e.g., Aubry et al. 2016;
Aubry and Igreja 2009; García-Rojas et al. 2021; Gómez de Soler et al.
2020; Herrero-Alonso, Fuertes-Prieto, and Neira-Campos 2020; Matias
2016; Ortega 2003; Pereira, Andrade, et al. 2016; Pereira et al. 2021,
2022; Ramacciotti et al. 2022; Sánchez de la Torre et al. 2023; Soto
2016; Rodríguez, Rodríguez, and Pelegrin 2011; Costa et al. 2022; Nocete
et al. 2005), and contributed to understanding how different groups
explored and managed the available lithic resources.

In this context, raw material analyses and the distinction between
local, regional and long-distance sources have been instrumental in
tracing extensive movements, contacts, and social networks throughout
the Upper Palaeolithic (UP) throughout Iberia. Noteworthy examples
include studies in northwestern Iberia (Galicia) during the Solutrean
period (Hermida, Rellán, and Rodríguez 2016), northern Iberia (Asturias)
during the Magdalenian occupations at Las Caldas (Corchón Rodríguez,
Martínez, and Tarriño 2016), northeastern Iberia (Catalunya) during the
Aurignacian and Gravettian periods at Arbreda cave (Marreiros et al.
2016; Ortega 2003), northern Portugal during the Gravettian/Solutrean
periods in the Côa Valley (Aubry et al. 2016, 2004), and inland central
Iberia (Guadalajara) during the Proto-Solutrean and Solutrean periods at
Peña Capón (Sánchez de la Torre et al. 2023). Ultimately, the
recognition of these connections has broadened our comprehension of
mobility and social organization during the UP in Iberia, illustrating a
vast interconnected territory.

However, in contrast to other areas of Iberia, and despite the potential
of raw material studies to shed light on the lifeways and organisational
structures of Palaeolithic hunter-gatherers, research on the procurement
and management of raw materials during the UP in southern and western
Iberia remains in its initial stages. Specifically, in southern
Portugal, even though there is considerable knowledge about the
technological organisation of Palaeolithic hunter-gatherers---largely
due to research conducted at the Vale Boi archaeological site (see e.g.,
Belmiro et al. 2021; Nuno Bicho et al. 2013; Cascalheira and Bicho 2013;
Marreiros et al. 2015; Cascalheira 2010; Horta, Cascalheira, and Bicho
2019)---raw material studies are notably infrequent and use mainly
macroscopic methodologies.

This situation is attributed to the absence of a detailed and
comprehensive characterization of regional lithic resources in the
Algarve region (southern Portugal) and the subsequent lack of a complete
reference collection produced by complementing characterization
methodologies, which is a critical component for conducting raw material
studies. Such efforts are particularly crucial for distinguishing
between local and non-local raw materials and for understanding
variations within the archaeological record (Pop 2015; Surovell 2009).

Recently, we have made significant strides by introducing such a
reference collection and establishing a framework for raw material
studies in the region, through a detailed macroscopic and petrographic
database of chert resources in the Algarve region (Belmiro, Terradas,
and Cascalheira 2023). This initiative paves the way for systematic
investigations into raw material procurement strategies during the Upper
Palaeolithic, especially concerning the Vale Boi site, where previous
raw material studies have already shown complex patterns of chert
procurement and management. These preliminary studies at Vale Boi,
employing only macroscopic methods, pointed to a predominant use of
local cherts (Nuno Bicho et al. 2013; Pereira, Bicho, et al. 2016;
Verissimo 2004), likely originating within a 16 km radius and reflecting
embedded procurement strategies (Pereira, Bicho, et al. 2016). To a
lesser extent, non-local cherts were also identified throughout the
various UP occupations (Nuno Bicho et al. 2013; Pereira, Bicho, et al.
2016; Verissimo 2004), pointing to extended sourcing networks, possibly
involving sourcing from other regions of southern Portugal and southern
Spain. The presence of chert from the Cretaceous formations of central
Portugal has also been suggested (Nuno Bicho et al. 2013).

These results highlight the potential for systematic raw material
studies at the site to shed light on the lithic resource procurement
strategies, mobility patterns, and social networks of the
hunter-gatherer groups of southwestern Iberia throughout the UP.

Vale Boi stands as a key archaeological site for this study since it is
currently the only site in southern Portugal with a long-term occupation
spanning most of the UP and allowing the exploration of raw material
procurement and mobility trends through time (Cascalheira et al. 2017).
It also holds a pivotal role in the study of UP adaptations on the
Iberian Peninsula, providing invaluable insights into the subsistence
and technological strategies of hunter-gatherer populations (Belmiro et
al. 2021; Nuno Bicho et al. 2013; Cascalheira et al. 2017; Cascalheira
2019; Horta, Cascalheira, and Bicho 2019; Manne et al. 2012; Marreiros
et al. 2016; Pereira, Bicho, et al. 2016).

Furthermore, previous studies have highlighted the role of Vale Boi in
understanding different patterns of territory exploitation and contact
with other regions of Iberia throughout the UP. During the Solutrean,
Vale Boi has been interpreted as an important connection point between
the core Solutrean areas of central Portugal and southern Spain, making
it a key site for identifying extended social networks that promoted the
exchange of information and culture (Cascalheira et al. 2017;
Cascalheira and Bicho 2013; Cascalheira 2013). On the contrary,
techno-typological studies from the Gravettian occupations of Vale Boi
instead show marked differences between regions and suggest a more
limited circulation of people and information in southern Iberia
(Marreiros and Bicho 2013).

Based on the notions that a) during the Solutrean Vale Boi served as a
connection point between different regions which may have promoted the
exchange of information and culture, including raw materials from
long-distances; and b) the limited circulation of people during the
Gravettian may have isolated groups occupying Vale Boi, making them
mostly reliant on the available local or regional resources, we propose
a set of hypotheses and expectations regarding chert procurement and
human mobility throughout the UP at Vale Boi. On one hand, it would be
expected that Solutrean occupations are characterized by the predominant
use of local raw materials but with a considerable percentage of cherts
obtained from long-distance sources, possibly through exchange with
different groups in Iberia. Conversely, it would be expected that
Gravettian occupations would be mostly composed of local cherts, with a
more limited presence of long-distance cherts.

To investigate the applicability of these patterns to the procurement
and circulation of lithic raw materials, and to examine the prevailing
hypotheses concerning UP mobility and social network exchanges, this
paper focuses on the following questions: 1) what types of chert were
the hunter-gatherers of Vale Boi using and where did they come from?; 2)
are there different patterns of chert use throughout time?; 3) and if
so, how can these be related to patterns of land use and the circulation
of people/ideas, as suggested through technological studies of the same
lithic assemblages? As a result, our study presents the first
comprehensive characterisation of chert raw materials, through
macroscopic and microscopic analysis, of UP occupations in southwestern
Iberia, derived from the Gravettian, Proto-Solutrean, and Solutrean
(c.~32-19 ka cal BP) occupations at Vale Boi.

\subsection{Site description}\label{site-description}

Vale Boi is located on the western coast of the Algarve region in
southern Portugal, within a small valley that extends southward to the
Atlantic coast, approximately 2 km away (Figure~\ref{fig-map} b). It is
bordered by limestone outcrops that form rock shelters facing west and
southwest (Nuno Bicho et al. 2007; Nuno Bicho, Cascalheira, and
Marreiros 2012; Nuno Bicho et al. 2013; Cascalheira and Bicho 2013;
Cascalheira et al. 2017; Manne et al. 2012; Manne and Bicho 2011). The
site extends for more than 10,000 sq meters along the slope of the
valley, through which three main areas were excavated between 2000 and
2019: Slope, Terrace, and Shelter. For this study, the chosen areas were
the Terrace and Shelter areas (see Section 2 Materials and Methods).

The Terrace encompasses occupations from the UP to the Early Neolithic,
resulting in the identification of eight main lithostratigraphic units.
The UP sequence includes several occupations attributed to the
Gravettian (levels 8 to 6) between c.~32 and 27 ka cal BP,
Proto-Solutrean (levels 5 to 4E) between c.~26 and 24 ka cal BP, and
Solutrean (level 4D, 4C, 4C, 4 and Lower 3) between c.~24 and 20 ka cal
BP (Belmiro et al. 2021; Cascalheira and Bicho 2013; Cascalheira et al.
2017).

The Shelter area shows four main lithostratigraphic units, with
Magdalenian, Solutrean and Gravettian occupations. From these, the
Solutrean levels show the most intensive occupation, with three,
well-preserved, archaeological horizons (layers C to A) and dated to
c.~24-22 ka cal BP (Cascalheira et al. 2013, 2017; Cascalheira 2010).
These occupation levels were identified under blocks of limestone, which
collapsed from the rockshelter ceiling (Cascalheira 2010).

Both these areas have been previously interpreted as seasonal
residential camps, repeatedly used for extended stays, due to the
abundance of lithic debitage, stone tools, heat-cracked rocks related to
grease rendering activities, large quantities of faunal remains (both
marine and terrestrial) and the presence of ornaments and portable art
(Manne et al. 2012). The analysis of lithic assemblages and retouched
frequency of lithic assemblages from the Terrace and Shelter areas also
corroborate this interpretation (Cascalheira 2010; Marreiros 2009;
Cascalheira et al. 2017). Cascalheira et al. (2017) show that, with the
exception of the Early Gravettian occupations of the Terrace area and
Magdalenian occupation of the Shelter, all other UP occupations are
composed of high-density assemblages with low degree of retouch,
correspondent to a residential base-camp occupation, for extended
periods of time.

\begin{figure}

\centering{

\includegraphics{figures/within-and-beyond/map_layout.png}

}

\caption{\label{fig-map}Iberian Peninsula map with locations and
geological formations mentioned throughout the paper. a) Detail of the
central Portugal region where the Cenomanian formations with chert
nodules are located. b) Detail of southwestern Portugal, with chert and
chalcedony bearing formations, outcrops, and location of Vale Boi. c)
Detail of eastern Algarve with chert-bearing formations. The figure was
produced using ArcGIS Pro 3.2.}

\end{figure}%

\subsection{Regional geological
background}\label{regional-geological-background}

The Algarve region is a complex territory, marked by several geomorphic
sub-regions and geological units that provide different lithological
resources. Two of the main geological units which characterize the
Algarve are: 1) the Baixo Alentejo Flyschoid Group, part of the south
Portuguese Zone (Fernandes et al. 2012; Oliveira 1984; Onézime et al.
2003), located in the north sector of the region, and characterized by
schists, greywackes and quartzites; and 2) the Algarve basin in the
south, which is characterized by limestones, dolomites and marls
(Terrinha et al. 2006). It is within these limestones and dolomitized
limestones from the Triassic to the Upper Jurassic where chalcedony and
chert nodules/beds can be found (Manuppella et al. 1987; Manuppella et
al. 2007; Ribeiro 2005; Rocha et al. 1979; Rocha et al. 1983; Terrinha
et al. 2006).

The availability of regional resources containing cherts and the
characterization of their lithologies has already been published by our
team (Belmiro, Terradas, and Cascalheira 2023), and we herein present
the summarized main findings. The results showed the existence of
abundant chert nodules and beds found in the Lower Jurassic (Carixian)
formation and Upper Jurassic (Kimmeridgian) formation, mostly
outcropping in the cliff and beach areas of westernmost Algarve, at
\textasciitilde20 km southwest from the site of Vale Boi, and with one
single inland outcrop located \textasciitilde8 km east from the site
(Figure~\ref{fig-map} b). The Lower Jurassic nodules and beds can be
found in abundant outcrops, but also frequently in sub-primary or
secondary deposition settings in proximity to the outcrops, possibly due
to natural erosion. However, in the case of the inland outcrop of
Ferrel, the outcrop is partially destroyed due to recent human action,
and chert can be found scattered as chunks or larger nodules. The Upper
Jurassic nodules can be found in sub-primary deposition and are easily
accessible, or in secondary deposition. The cherts from the previously
mentioned formations are are frequently opaque with yellow, grey, purple
or red colours, occasionally with bands or laminations, and frequently
with visible sponge spicules. Although abundant, these chert nodules are
frequently small (\textasciitilde4-10 cm in length), with the largest
nodules being \textasciitilde20 cm, and frequently characterized by high
levels of dolomitization and fractures. Petrographically, they are
composed mainly of microcrystalline quartz, with wackestone or packstone
textures, small percentages of macrocrystalline quartz, fibrous
chalcedony and allochems composed of oxides and poorly preserved
bioclasts. When identifiable, the bioclasts are Echinoderms,
Radiolarians, Sponge spicules and Bivalve shells. The Upper Jurassic
cherts of Mareta show similar petrographic characteristics, but with
abundant Calcispheres which are not always apparent under the
stereomicroscope.

Chert sources were also identified in two formations from the eastern
area of the Algarve (Figure~\ref{fig-map} c). The Middle Jurassic Malhão
formation chert outcrops were identified at around 80-100 km east of the
site and characterized by chert nodules with similar macroscopic
features to those found in western Algarve. Petrographically they are
composed mainly of microcrystalline quartz and characterized by
wackestone textures, while dolomite, chalcedony and macrocrystalline
quartz are also present in small percentages (\textless10\%). Allochems
are iron oxides, and albeit poorly preserved, several bioclasts can be
identified: Sponge spicules, Radiolarians, Ostracods, Echinoderms,
Calcispheres and possibly Tentaculites. The Upper Jurassic Jordana
formation chert outcrops were identified at around 90 km east of Vale
Boi and characterized by chert nodules with significantly different
macroscopic features to those of other formations, with frequently small
sizes (\textasciitilde5 cm), difficult removal from the parent rock and
unidentified secondary deposits. Petrographically they are composed
mostly of microcrystalline quartz, with wackestone to packstone
textures, and small percentages of fibrous chalcedony and dolomite
(\textless10\%). They show frequent iron oxides and very frequent
bioclasts, albeit often poorly preserved and unidentifiable; identified
fossils are Calcispheres, Bivalve shells, Sponge spicules, Ostracods,
Echinoderms and Gastropods.

Chalcedony can be found \textasciitilde10 km west of Vale Boi
(Figure~\ref{fig-map} b), within the marl-carbonated Triassic formations
as lenses or nodules and was formed through hydrothermal processes (Nuno
Bicho et al. 2013). Previous studies also note the presence of
chalcedony (with frequent presence of cortex) in the immediate vicinity
of the site (Pereira, Bicho, et al. 2016; Gibaja Bao and Ferreira Bicho
2013). The chalcedony frequently has fractures and fissures and is
characterized by a fibrous quartz structure, without inclusions or
fossils.

\section{Materials and methods}\label{materials-and-methods}

To characterize the siliceous raw materials used during the UP
occupations at Vale Boi and to identify their geological sources, we
focused on chalcedony and chert archaeological materials from the
Terrace and Shelter areas. These two areas were chosen since both have
detailed spatial information, well-defined chronological sequences, and
previous studies providing lithic technology, faunal, and
geoarchaeological data. For this, we focused on layers that spanned the
UP occupations, characterized by good preservation, a significant number
of lithic finds and with radiocarbon dating (as mentioned in section
Site description).

In this study, the terms chert and chalcedony follow the definitions
provided by Luedtke (1992). We use the term chert to refer to
sedimentary rocks composed primarily of microcrystalline quartz, which
includes other possible varieties such as jasper, agate, or chalcedony.
The term chalcedony is used in its petrographic sense and refers to
cherts with a fibrous quartz structure. Since this definition of
chalcedony implies a destructive method which cannot be applied to all
archaeological chalcedony samples, we use the term based on the
confirmed correlation between the macroscopic appearance of chalcedony
and its petrographic characteristics, obtained through geological and
archaeological thin sections.

Chert was chosen for this study as it is one of the three main raw
materials used at the archaeological site of Vale Boi alongside quartz
and greywacke. Other raw materials such as chalcedony, dolerite, and
schists have also been identified (Belmiro et al. 2021; Cascalheira
2010; Pereira, Bicho, et al. 2016).

Using the field database with all individually coordinated lithic
artefacts (all pieces larger than \textasciitilde2 cm; (Belmiro 2020)),
we compared the incidence of chert and chalcedony in opposition to the
remaining lithics of other raw materials. Figure~\ref{fig-rm-comparison}
shows that for the selected levels of the Terrace area
(Figure~\ref{fig-rm-comparison} a), chert represents only
\textasciitilde15-20\% of the lithic assemblage, while in the Shelter
area (Figure~\ref{fig-rm-comparison} b), chert represents
\textasciitilde35-50\% of the lithic assemblage from the Solutrean
layers A, B, and C. Based on previous lithic studies in the Terrace and
Shelter areas, the remaining 80\% of other raw materials are mostly
composed of quartz and greywacke (Belmiro et al. 2021; Cascalheira 2010;
Marreiros 2009).

Although systematic raw material research has not been conducted on
these materials, according to the regional geological context, they have
been interpreted as local and readily accessible in the vicinity of the
site (Pereira, Bicho, et al. 2016). Frequently, the quartz and greywacke
assemblages include large amounts of shattered and unknapped chunks or
slabs, and have been interpreted in great part as functionally
specialised in activities related to the fragmentation of bones in the
case of quartz and greywack (Cascalheira et al. 2017), or in grease
rendering in the case of quartz (Manne et al. 2012). Previous studies
from the Terrace area have also shown that when removing shatter and
chunks, the percentages of chert become much more representative when
compared to greywacke and quartz (Belmiro et al. 2021; Belmiro 2020).
Based on this, despite the smaller amounts of individually plotted chert
artefacts in the field database, chert continues to be one of the main
raw materials to produce lithic stone tools, with complete knapping
sequences and formal toolkits. The importance of chalcedony can also be
seen in its use to produce formal tool kits during specific
technocomplexes at the site, such as a Vale Comprido point, the
fossil-director for the Proto-Solutrean in western Portugal (Belmiro et
al. 2021) or a laurel leaf preform in the Solutrean layers (Gibaja Bao
and Ferreira Bicho 2013).

\begin{figure}

\centering{

\includegraphics{article2_files/figure-pdf/fig-rm-comparison-1.pdf}

}

\caption{\label{fig-rm-comparison}Comparison of chert, chalcedony, and
other raw materials piece plotted in the Terrace (a) and Shelter (b)
areas of Vale Boi.}

\end{figure}%

We employed both macroscopic and petrographic methods to achieve a more
comprehensive characterization of the materials, aiming to offset the
inherent limitations of each method (Luedtke 1992).

Macroscopic analysis, while revealing several limitations related to the
absence of quantitative variables (Bustillo et al. 2009), proved to be
advantageous given the extensive sample size of chalcedony and chert
lithics under examination (n=4458). The distinct features of various
local cherts facilitate their differentiation from cherts of
neighbouring origins. Consequently, a macroscopic approach offered a
non-destructive, efficient, and cost-effective method for characterizing
all materials (Bustillo et al. 2009; Tarriño and Terradas 2013).
Therefore, we implemented an initial phase of macroscopic analysis,
subsequently augmented by petrographic study to diminish the
subjectivity of the analysis and enhance the material characterization
with petrographic data. The synergistic application of these two methods
has been previously validated in the characterization of Algarve region
cherts, following recent geological survey efforts (Belmiro, Terradas,
and Cascalheira 2023). Hence, an extensive macroscopic and petrographic
collection was already available for comparative analysis with
archaeological materials, employing consistent methodologies. This
reference collection is the Lusolit lithotheque, hosted physically at
ICArEHB (University of Algarve, Faro) and digitally at
www.lusolit.icarehb.com, and aided in the characterization and source
attribution of cherts.

Hand samples were used throughout the macroscopic characterization for
comparison with the archaeological materials, and the archaeological
thin sections were compared with the previously studied geological chert
thin sections. To identify the source of other lithologies which are not
congruent with the regional cherts and identify possible long-distance
chert procurement, we used hand samples and thin sections of cherts from
central Portugal also available in the Lusolit lithotheque. The
lithotheque from the Unit of Geoarchaeology and Archaeometry Applied to
Historic Artistic and Monumental Heritage from the University of Cadiz
(UCA) was also visited to understand the macroscopic variability of the
local and regional cherts of the lower Guadalquivir basin, within the
Atlantic strip of southwestern Spain (present-day provinces of Huelva
and Cádiz).

Following previous studies, we classify the distance of chert sources
from the site as local (1-30 km from the site), regional (30-120 km),
and long-distance (\textgreater120 km) (Herrero-Alonso, Fuertes-Prieto,
and Neira-Campos 2020; Tarriño, Elorrieta, and García-Rojas 2015).

The macroscopic analysis of the archaeological materials was divided
into two steps. The first step included a preliminary characterization
using a hand lens of 10x magnification. This step allowed us to create
types and subtypes based on similar macroscopic characteristics
frequently acknowledged to be useful to describe and discern between
different types of chert; these included characteristics such as colour,
translucency, and feel (Luedtke 1992)-the complete data dictionary for
the recorded variables can be found in the Supplementary Information
(Online Resource 1). In the second step, we used a Nikon SMZ25
stereomicroscope to observe each sample in further detail, which allowed
us to better characterize the artefacts (especially regarding inclusions
and alterations), and the previously established types - the data
dictionary used for the individual samples can also be found in the
Supplementary Information (Online Resource 1). Given the inherent
macroscopic variability of the geological samples and archaeological
specimens, as well as the number of samples, the database used for the
individual analysis was simplified and focused on collecting data
related to weight, cortex, and alterations. This first step was applied
to a total of 4458 artefacts: a sample of 3627 chert artefacts from the
UP sequence of the Terrace area (levels 7 to 4) and 831 chert artefacts
from the Solutrean sequence of the Shelter area (layers A, B, C; all
samples without an attributed layer on the database were excluded from
the analysis).

While the advantages of employing a macroscopic methodology for
analysing the cherts of Vale Boi are notable, a limitation of this
approach is its inability to unequivocally characterize and categorize
artefacts that exhibit surface alterations typical of an open-air site,
including those related to fire, since weathered and altered surface may
destroy visible structures and even alter the chemical composition
(Delluniversità et al. 2019). Other studies frequently exclude samples
that display intense fire damage or several surface alterations
(Delluniversità et al. 2019; Gómez de Soler et al. 2020; Soto 2016). For
the present study, whenever artefacts displayed extensive alterations
and could not be unequivocally characterized
(Figure~\ref{fig-terrace-alterations}), and given the inexistence of a
reference collection for local altered cherts, they were analysed using
the individual sample dataset (Online Resource 1) but grouped in a
category of indeterminate cherts (INDET). This category included key
alterations such as heat/fire-related alterations which clearly impacted
the macroscopic characteristics of the samples, fully changing its
color, luster and texture, and covering them in crazings (e.g.,
Figure~\ref{fig-terrace-alterations} a). Similarly, it also included
alteration rinds that extended through most of the sample, impeding the
identification of its original colour, texture, translucency, inclusions
and fossils (e.g., Figure~\ref{fig-terrace-alterations} b) and the
presence of extended pitting. However, whenever surface alterations were
present in only a limited area of the sample leaving recognizable
macroscopic features and fossils, or when key elements were
distinguishable to make a characterization (e.g., reddening in the Type
2 cherts, Online Resource 2, figure s5), the samples were not collapsed
into the INDET category.

As seen in Table~\ref{tbl-indet-freq}, level 4 shows the highest
percentages of indeterminate artefacts, corresponding to
\textasciitilde40\% of the total sample (n=1037). This is due to the
frequent alterations of the artifacts, especially related to patinas and
alteration rinds, pits, and fire/heat alterations. For all other levels
in the Terrace, indeterminate represents on average \textasciitilde28\%
of the chert artefacts, while in the Shelter the frequency of INDET is
slightly lower (\textasciitilde20-25\%).

\begin{figure}

\centering{

\includegraphics{figures/within-and-beyond/terrace_alterations.png}

}

\caption{\label{fig-terrace-alterations}Samples with extensive
alterations which prevented the identification of chert type. a)
artefact with cracks, white patina and reddening covering 100\% of the
surface; b) artefact with surface alterations, displaying a portion with
patina (brown) and a portion without patina (grey), cleaned with
Hydrochloric acid.}

\end{figure}%

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0828}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0947}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0828}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0828}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0828}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0828}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0828}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0828}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0828}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0769}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0828}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 22\tabcolsep) * \real{0.0828}}@{}}

\caption{\label{tbl-indet-freq}Number and percentage of grouped and
indeterminate (INDET) artefacts by level in the Terrace (levels 4 to 7)
and Shelter (A to C) areas.}

\tabularnewline

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Group
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
4
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
4C
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
4D
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
4E
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
4E/5
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
5
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
6
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
7
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
A
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
B
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
C
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Grouped & 619 (59.69\%) & 287 (71.75\%) & 287 (66.74\%) & 332 (73.45\%)
& 155 (73.11\%) & 211 (71.04\%) & 434 (73.06\%) & 149 (72.68\%) & 72
(72.73\%) & 212 (69.28\%) & 349 (81.92\%) \\
Indeterminate & 418 (40.31\%) & 113 (28.25\%) & 143 (33.26\%) & 120
(26.55\%) & 57 (26.89\%) & 86 (28.96\%) & 160 (26.94\%) & 56 (27.32\%) &
27 (27.27\%) & 94 (30.72\%) & 77 (18.08\%) \\
Total & 1,037 (100.00\%) & 400 (100.00\%) & 430 (100.00\%) & 452
(100.00\%) & 212 (100.00\%) & 297 (100.00\%) & 594 (100.00\%) & 205
(100.00\%) & 99 (100.00\%) & 306 (100.00\%) & 426 (100.00\%) \\

\end{longtable}

As previously mentioned, the macroscopic study was followed by a
petrographic approach. To better characterize the different chert types
identified in the macroscopic analysis and aid in minimizing the
inherent caveats of macroscopic methods, 22 thin sections were prepared
at the Thin Section Lab (Toul, France). The samples chosen for this
analysis include at least one sample from each type present throughout
the stratigraphy of the Terrace area, from different archaeological
levels (Table~\ref{tbl-TS-list}). For types that showed high macroscopic
variability (e.g.~translucency and colour variation or differentiated
bioclast facies), thin sections from sub-types were analysed to collect
petrographic data that could reflect this variability. Petrographic thin
sections were analysed by polarized light microscopy (Nikon LV100ND) to
determine the mineral composition, textural characteristics, allochems,
and bioclasts, as well as any other alterations.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.0808}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1010}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.0606}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.7576}}@{}}

\caption{\label{tbl-TS-list}List of archaeological thin sections, with
the summarised description of observed macroscopic variability and
sub-types.}

\tabularnewline

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Sample ID
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Macroscopic variability/sub-type description.
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Type 1 & I19-1615 & 4D & - \\
Type 2 & I19-2835 & 5 & - \\
Type 2 & H18-1938 & 4D & Type 2 sample with surface alterations. \\
Type 2 & H21-3095 & 4C & Type 2 multi-coloured banding sub-type with
concentration of bioclasts. \\
Type 2 & H21-4234 & 5 & Type 2 multi-coloured banding sub-type without
concentration of bioclasts. \\
Type 2 & H19-2426 & 4D & Type 2 sample with reddening and possible
heat-related alterations. \\
Type 2 & I19-2226 & 4E & Type 2 sample with large iron oxides and rare
sponge spicules. \\
Type 3 & I20-3160 & 5 & - \\
Type 4 & H20-2441 & 4E & - \\
Type 5 & I21-3252 & 4E & - \\
Type 5 & H18-2708 & 5 & Type 5 banding sub-type. \\
Type 6 & I19-3350 & 6 & Type 6 sample with light white patination. \\
Type 6 & I20-3689 & 6 & - \\
Type 6 & J18-778 & 5 & Type 6 with thick white patination. \\
Type 6 & I20-3951 & 6 & Type 6 light brown with opaque white exterior
sub-type. \\
Type 6 & H19-4074 & 6 & Type 6 with lamination and white broad mottling
sub-type. \\
Type 7 & L19-64 & 5 & Type 7 with sponge spicule (01a) and peloidal
(01b) facies. \\
Type 7 & J18-1264 & 8 & Type 7 with peloidal/oolitic (04) facies. \\
Type 8 & H19-2924 & 4E & - \\
Type 9 & I21-2966 & 4C & - \\
Type 10 & H19-4216 & 6 & - \\
INDET & H20-4166 & 6 & Sample with thick white patination. \\

\end{longtable}

The complete petrographic analysis for all thin sections studied in this
paper (including macroscopic and microscopic figures) can be found at
our online research compendium (https://doi.org/10.17605/OSF.IO/DBFT2).
A detailed macroscopic and petrographic description and figures for each
type and sub-type of chert and chalcedony (Online Resource 2) and the
data dictionary used for the petrographic analysis (Online Resource 1)
can be found in Supplementary Information. The complete set of images of
the geological thin sections can also be consulted in the LusoLit online
database.

The entirety of the R code used for the analysis, datasets, and visual
representations contained in this paper can be accessed through our
online research compendium. We used the rrtools package by Marwick et
al. (2018) to create a research compendium and write a reproducible
journal article. The provided files include the complete set of raw data
used in the analysis, along with a custom R project (Wickham 2015)
containing the code required to generate all tables and figures. To
enable maximum reuse, the code is made available under the MIT license,
data under CC‐0, and figures under CC‐BY (additional details can be
found in Marwick 2017).

\section{Results}\label{results}

\subsection{Characterization and source
attribution}\label{characterization-and-source-attribution}

Through the macroscopic and petrographic analysis, we identified 10 main
chert/chalcedony types, present throughout the stratigraphy in the
Terrace and Shelter areas (a summarised characterization table can be
found in the Supplementary Information, Online Resource 3). A distinct
type (Type 11) was identified exclusively in the Shelter area, with only
a limited presence in the Terrace (n=2), and 16 varieties, each
consisting of one to three samples, were identified across various
levels of both the Terrace and Shelter. The latter were clustered into a
category named Trace Lithotypes (TL). Through the comparison with the
reference collections, we identified cherts congruent with the local
sources and cherts that are likely not local due to significant
differences in appearance, fossil content, and petrography when compared
to the cherts available in the geological setting located within a 30 km
radius of Vale Boi. Table~\ref{tbl-type-source} shows the different
types of identified cherts and possible sources whenever attribution was
possible.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1358}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.6049}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2593}}@{}}

\caption{\label{tbl-type-source}Chert types identified during the
analysis, and potential sources. The distance column refers to the
distance of the potential source from Vale Boi, as the crow flies. Only
the TL samples with possible identified sources were included in the
table.}

\tabularnewline

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Chert type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Potential source
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Distance from source
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Type 1 & Triassic volcanic formations & \textless10 km \\
Type 2-5 & Lower Carixian formations, southwestern Portugal &
\textless20 km \\
Type 6 & Upper Cretaceous formations, central Portugal &
\textasciitilde215-250 km \\
Type 7 & Betic Systems, southern Spain & \textasciitilde250 km \\
Type 8 & - & - \\
Type 9 & Betic Systems, southern Spain & \textasciitilde250 km \\
Type 10 & - & - \\
Type 11 & - & - \\
TL01 & South Portuguese Zone, southern/central Portugal &
\textgreater100 km \\
TL10-11 & Upper Cretaceous formations, central Portugal &
\textasciitilde215-250 km \\
TL15-16 & Betic Systems, southern Spain & \textasciitilde250 km \\

\end{longtable}

\subsubsection{Cherts congruent with local sources (Types
1-5)}\label{cherts-congruent-with-local-sources-types-1-5}

The raw materials which are congruent with local sources are the
chalcedony and massive, micro-cryptocrystalline cherts
(Figure~\ref{fig-local-types}). The chalcedony (Type 1;
Figure~\ref{fig-local-types} a) is identified as a massive, fibrous
variety without allochems (Figure~\ref{fig-local-types} a2). It
represents 6.6\% of the overall Terrace cherts (excluding INDET) and
2.6\% in the Shelter. It presents a colourless and translucent
appearance, frequently exhibiting partial or complete white patination
(Figure~\ref{fig-local-types} a1). This type is marked by extensive
crystallizations, irregularities, and fractures, coupled with
post-depositional alterations that challenge the identification and
preservation of the cortex. Both macroscopic and petrographic analyses
confirm its congruence with hydrothermal chalcedony located within
lenses in Triassic volcanic formations, in the known outcrops of Monte
do Tolo, Taboal and Cerro Velho (\textless10 km west of Vale Boi, as the
crow flies; Figure~\ref{fig-map} b), which are similarly marked by
irregularities and fractures.

The cherts are massive micro-cryptocrystalline quartz from a marine
environment (Figure~\ref{fig-local-types} b-h). They represent
\textasciitilde60\% of the overall Terrace cherts and
\textasciitilde80\% of the chert in the Shelter. Macroscopically, these
types show variability which resulted in their separation in four
different macroscopic types (Type 2 to 5). In general, these are opaque
to sub-translucent cherts, with colours varying from brown, yellow, pale
red, and grey, with the frequent presence of spicules and iron oxides.
The differences are in their colours, nodule morphology, frequency, and
variety of bioclasts (Type 2 including triaxon sponge spicules and
echinoderm spines), something already identified in the geological
samples (Belmiro, Terradas, and Cascalheira 2023). As seen in table
Table~\ref{tbl-cortex_tab}, whenever present, the cortex seems to show a
mixture of irregular cortex and rounded cortex, both from an outcrop
source. Petrographically, however, these cherts are homogeneous, except
for a small number of artefacts with a banded variety within Type 5.
These cherts are characterized by wackestone textures, with accessory
amounts of macrocrystalline quartz and fibrous chalcedony, frequently
found replacing fossils and occasionally dolomite. Type 5 shows higher
percentages of micrite/sparite. In all samples, fossils range from
common to very frequent, albeit frequently unidentifiable. When
identifiable in thin section, they are sponge spicules
(Figure~\ref{fig-local-types} b2, c2, h2), with the rare occurrence of
possible foraminifera (Type 2 and Type 5) and ostracod
(Figure~\ref{fig-local-types} h3) in a sample from Type 5. Porosity
varies between 10\% and 1\%, of vug type. Despite the petrographic
homogeneity between local types, this macroscopic separation was
maintained for its possible usefulness in identifying preferred types of
nodules or macroscopic varieties through time and for future correlation
with technological data.

\begin{table}

\caption{\label{tbl-cortex_tab}Frequency of cortex type and thickness by
raw material from the Terrace area, characterized during the individual
analysis of the lithic assemblage.}

\centering{

\fontsize{12.0pt}{14.4pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lccccccccccc}
\toprule
\textbf{Variable} & \textbf{TYPE 1}  N = 8\textsuperscript{\textit{1}} & \textbf{TYPE 2}  N = 267\textsuperscript{\textit{1}} & \textbf{TYPE 3}  N = 138\textsuperscript{\textit{1}} & \textbf{TYPE 4}  N = 112\textsuperscript{\textit{1}} & \textbf{TYPE 5}  N = 134\textsuperscript{\textit{1}} & \textbf{TYPE 6}  N = 163\textsuperscript{\textit{1}} & \textbf{TYPE 7}  N = 108\textsuperscript{\textit{1}} & \textbf{TYPE 8}  N = 19\textsuperscript{\textit{1}} & \textbf{TYPE 9}  N = 18\textsuperscript{\textit{1}} & \textbf{TYPE 10}  N = 3\textsuperscript{\textit{1}} & \textbf{TL}  N = 3\textsuperscript{\textit{1}} \\ 
\midrule\addlinespace[2.5pt]
{\bfseries Cortex type} &  &  &  &  &  &  &  &  &  &  &  \\ 
    Irregular & 1 (13\%) & 74 (28\%) & 52 (38\%) & 50 (45\%) & 31 (23\%) & 20 (12\%) & 30 (28\%) & 2 (11\%) & 0 (0\%) & 0 (0\%) & 0 (0\%) \\ 
    Rounded & 1 (13\%) & 118 (44\%) & 53 (38\%) & 35 (31\%) & 69 (51\%) & 116 (71\%) & 46 (43\%) & 14 (74\%) & 14 (78\%) & 1 (33\%) & 3 (100\%) \\ 
    Pebble & 0 (0\%) & 4 (1.5\%) & 2 (1.4\%) & 0 (0\%) & 2 (1.5\%) & 6 (3.7\%) & 1 (0.9\%) & 0 (0\%) & 0 (0\%) & 0 (0\%) & 0 (0\%) \\ 
    Indeterminate & 6 (75\%) & 71 (27\%) & 31 (22\%) & 27 (24\%) & 32 (24\%) & 21 (13\%) & 31 (29\%) & 3 (16\%) & 4 (22\%) & 2 (67\%) & 0 (0\%) \\ 
{\bfseries Thickness} &  &  &  &  &  &  &  &  &  &  &  \\ 
    Thin & 7 (88\%) & 121 (45\%) & 49 (36\%) & 24 (21\%) & 105 (78\%) & 151 (93\%) & 37 (34\%) & 13 (68\%) & 13 (72\%) & 0 (0\%) & 3 (100\%) \\ 
    Medium & 1 (13\%) & 122 (46\%) & 77 (56\%) & 53 (47\%) & 23 (17\%) & 10 (6.1\%) & 51 (47\%) & 5 (26\%) & 3 (17\%) & 3 (100\%) & 0 (0\%) \\ 
    Thick & 0 (0\%) & 24 (9.0\%) & 12 (8.7\%) & 35 (31\%) & 6 (4.5\%) & 2 (1.2\%) & 20 (19\%) & 1 (5.3\%) & 2 (11\%) & 0 (0\%) & 0 (0\%) \\ 
\bottomrule
\end{tabular*}
\begin{minipage}{\linewidth}
\textsuperscript{\textit{1}}n (\%)\\
\end{minipage}

}

\end{table}%

Based on the previously described macroscopic and petrographic
characteristics of the Lower Jurassic cherts of the Algarve (Belmiro,
Terradas, and Cascalheira 2023), we interpret Types 2, 3, 4, and 5 to be
local, as they are congruent with the described characteristics
representative of cherts from the Lower Jurassic found in the known
outcrops Cabo S. Vicente, Foz dos Fornos, Ponta dos Altos, Praia do
Belixe and Ferrel (Carixian; \textless20 km from Vale Boi, as the crow
flies), possibly Upper Jurassic Kimmeridgian outcrops of Mareta
(\textasciitilde15 km from Vale Boi, as the crow flies) and Middle
Jurassic Malhão formations (80-100 km east of Vale Boi, as the crow
flies) of the Algarve basin. Types 2 and 3 are congruent with the
frequently available brown/yellow and red nodules and beds
(Figure~\ref{fig-local-types} b-d), ranging from uncommon to very
frequent fossils. These can be found in all previously mentioned
outcrops. The macroscopic similarities of Types 4 and 5 are more limited
(Figure~\ref{fig-local-types} e-h), as they show similarities to
specific nodules, rarely found in the known outcrops, and mostly located
in the inland Lower Jurassic outcrop of Ferrel (Figure~\ref{fig-map} b).
As such, we interpret these cherts as local, but possibly from nodules
and inland outcrops that are no longer frequently available or visible
in the landscape. This is especially relevant since the inland outcrops
of Ferrel are currently mostly destroyed and can be found in poor
preservation conditions, due to recent human intervention and
construction work, and with small, broken nodules in the surface,
frequently altered (Belmiro, Terradas, and Cascalheira 2023).

\begin{figure}

\centering{

\includegraphics{figures/within-and-beyond/local_grid.png}

}

\caption{\label{fig-local-types}Variability of local chert types and
main petrographic characteristics. a) Type 1 macroscopic and microscopic
views: a1) macroscopic view with a stereomicroscope of white patina and
fractures; a2) microscopic view of thin section (XPL), with visible
fibrous chalcedony composition and without allochems. b) Type 3
macroscopic and microscopic views: b1) macroscopic view with a
stereomicroscope of iron oxides and unidentifiable fossils; b2)
microscopic view of thin section (XPL), with visible unidentifiable
fossils and iron oxides. c-d) Type 2 macroscopic and microscopic views
of multicoloured (c) and single-colour (d) varieties. c1) macroscopic
view with a stereomicroscope of the multicoloured bands with iron oxides
and unidentifiable fossils; d1) macroscopic view with a stereomicroscope
of iron oxides and unidentifiable fossils; c2) microscopic view of thin
section (PPL), with visible unidentifiable fossils, sponge spicules and
iron oxides; d2) microscopic view of thin section (XPL), with visible
unidentifiable fossils and iron oxides. e-f) Type 4 macroscopic and
microscopic views: e) macroscopic view with a stereomicroscope of
unidentifiable fossils and white splotches; f1) macroscopic view with a
stereomicroscope of iron oxides and unidentifiable fossils within white
and yellow splotches; f2) microscopic view of thin section (XPL), with
visible unidentifiable fossils and iron oxides; f3) microscopic view of
thin section (PPL) with a concentration of fossils (possibly sponge
spicules) and a fracture. g-h) Type 5 macroscopic and microscopic views:
g) macroscopic view with a stereomicroscope; h1) macroscopic view with a
stereomicroscope with visible unidentifiable fossils; h2) microscopic
view of thin section (PPL), with visible unidentifiable fossils, sponge
spicules and iron oxides; h3) microscopic view of thin section (XPL)
with small unidentifiable fossils, iron oxides and a possible ostracod
replaced by fibrous chalcedony and macrocrystalline quartz.}

\end{figure}%

\subsubsection{Cherts not congruent with local sources (Types 6-11 and
TL)}\label{cherts-not-congruent-with-local-sources-types-6-11-and-tl}

We identified six chert types with macroscopic and petrographic
characteristics that are not congruent with the chert available in known
outcrops in western Algarve (Figure~\ref{fig-distance-types}).

\begin{figure}

\centering{

\includegraphics{figures/within-and-beyond/nonlocal_grid.png}

}

\caption{\label{fig-distance-types}Variability of non-local types and
main petrographic characteristics. a-b) Type 6 macroscopic and
microscopic views: a1) macroscopic view with a stereomicroscope with a
small degree of white patina and no visible fossils; b1) macroscopic
view with a stereomicroscope of a thick alteration rind but translucent
interior where some unidentifiable fossils are visible; a2) microscopic
view of thin section (PPL) with a mudstone texture; b2) microscopic of
thin section (XPL) with small unidentifiable fossils and a laminar
microstructure. c-d) Type 7 macroscopic and microscopic views of the
peloidal facies (c) and oolitic facies (d): c1) macroscopic view with a
stereomicroscope of peloids and sponge spicules; d1) macroscopic view
with a stereomicroscope of poorly preserved ooids; c2) microscopic view
of thin section (PPL) with packstone texture composed mainly of peloids;
d2) microscopic view of thin section (PPL) with packstone texture
composed mainly of ooids. e-f) Type 8 macroscopic and microscopic views,
with foraminifera represented by the white and black arrows: e)
macroscopic view with a stereomicroscope with abundant unidentifiable
fossils and foraminifera; f1) macroscopic view with a stereomicroscope
with abundant unidentifiable fossils but no visible foraminifera; f2)
microscopic view of thin section (PPL) with packstone texture composed
of unidentifiable fossils and foraminifera which are not visible
macroscopically; f3) microscopic view of thin section (PPL) with
packstone texture composed of unidentifiable fossils and foraminifera
which not are visible macroscopically. g) Type 9 macroscopic and
microscopic views: g1) macroscopic view with a stereomicroscope of
unidentifiable fossils and two different fabrics; g2) microscopic view
of thin section (XPL) with packstone texture composed of concentrations
of unidentifiable fossils and opaques which create a banded
microstructure. h) Type 10 macroscopic and microscopic views: h1)
macroscopic view with a stereomicroscope of ooids and fractures filled
with chalcedony; h2) microscopic view of thin section (PPL) with
packstone texture composed of ooids with different levels of
preservation.}

\end{figure}%

Type 6 (Figure~\ref{fig-distance-types} a-b) corresponds to a massive
cryptocrystalline quartz with a mudstone texture, from a marine
environment. It represents \textasciitilde18\% of the overall Terrace
cherts and \textasciitilde12\% of the chert in the Shelter. This chert
is translucent and shows a variety of colours, ranging from brown, red,
and grey, which seem to vary in function of the artefact's thickness. In
unaltered samples, visible bioclasts are rare. The cortex is mostly
rounded, and rarely pebble-like, ranging from thin to medium thickness
(Table~\ref{tbl-cortex_tab}). Two sub-types were individualized based on
macroscopic characteristics such as colour, translucency variation, and
patterns, although they may simply reflect the use of specific nodules
from within Type 6's variability. Five thin sections were produced to
better characterize the macroscopic variability of this type and
sub-types. The petrographic analysis shows the samples are characterized
by a massive microstructure, mainly composed of cryptocrystalline quartz
(between 95-97\% in the samples), with accessory fibrous chalcedony
found replacing fossils or filling fractures
(Figure~\ref{fig-distance-types} a2-b2), and in some samples
macrocrystalline quartz, microcrystalline quartz, and micrite/sparite.
Allochems are opaques, iron oxides, and bioclasts. Bioclasts are
unidentifiable and poorly preserved, with rare identifiable fossils such
as sponge spicules, replaced by fibrous chalcedony and rarely by
microcrystalline quartz. Porosity is present between \textless1-1\%, of
vug-type when identifiable.

The similarities between chert artefacts at the UP occupations of Vale
Boi and the chert from the Rio Maior region (central Portugal;
Figure~\ref{fig-map} a) have been previously suggested although without
systematic studies to ascertain the attribution (Nuno Bicho et al. 2013;
Pereira, Bicho, et al. 2016). These Cretaceous cherts, mostly found in
secondary deposition settings, are described as frequently translucent,
with geodes, colours ranging from yellow to red or grey, and
mineralogically homogeneous between them. Petrographic studies from
central Portugal highlight the presence of iron oxide accumulations and
rare fossil ghosts or frequently difficult to see in thin section
(Matias 2016). When identified, sponge spicules and possible rare
foraminifera have been identified (Matias 2012).

Studies of Upper Cretaceous cherts located in the Lisbon area describe
cherts with similar macroscopic characteristics (Figure~\ref{fig-map}
a), with micro-cryptocrystalline quartz mudstones with bioclasts such as
sponge spicules or bivalve fragments, or wackestone/packstone textures
with abundant bioclasts, including ostracods and different types of
foraminifers (Jordão and Pimentel 2022). These cherts have also been
reported as found in highly variable cobble morphologies and macroscopic
traits, which is also observable through the reference samples from
central Portugal hosted at LusoLit. Similar to the archaeological thin
sections from Type 6 and congruent with other studies (Jordão and
Pimentel 2022; Matias 2016), the thin sections produced from the
Cretaceous cherts hosted at Lusolit show a massive, cryptocrystalline
mudstone structure, with uncommon fossils (ghosts), iron oxides and low
porosity (\textless1\%). From these Cretaceous cherts from central
Portugal, only the dark grey nodules (sample RT231, with petrographic
descriptions available in our online research compendium) showed
identifiable fossils which were rare ostracods.

Based on the macroscopic and petrographic similarities to the geological
references from central Portugal and congruence with previous studies
from this area, we interpreted Type 6 (and sub-types) as belonging to
the macroscopically variable Cretaceous cherts (Upper Cenomanian
formations) from central Portugal, between \textasciitilde215-250 km (as
the crow flies) north of Vale Boi (Figure~\ref{fig-map} a).

Type 7 (Figure~\ref{fig-distance-types} c-d) is a massive
micro-cryptocrystalline, peloidal/oolitic packstone, from a marine
environment and possibly a high-energy, shallow depositional
environment, due to the type, sorting, and preservation conditions of
the ooids. In the terrace, it represents 10\% of the overall Terrace
cherts and less than 1\% of the chert in the Shelter. It is
characterized by white and grey colours. The cortex is from an outcrop,
rounded and in lesser frequencies (\textless30\%) irregular, and varying
between medium thickness to thick (Table~\ref{tbl-cortex_tab}). This
type shows several facies, with different types and concentrations of
allochems and bioclasts, such as peloids
(Figure~\ref{fig-distance-types} c1), ooids
(Figure~\ref{fig-distance-types} d1), and sponge spicules. A detailed
scheme of the facies can be found in the Supplementary Information
(Online Resource 2). The petrography results show this chert has a
massive microstructure and packstone texture
(Figure~\ref{fig-distance-types} c2-d2). It is composed mainly of
micro-cryptocrystalline quartz (80\%), fibrous chalcedony (10\%), and
micrite (10\%). Allochems are opaques, iron oxides, ooids, and peloids.
The ooids are poorly preserved and only uncommonly show concentric
lamellae in plane-polarized light. Unlike macroscopical observations,
bioclasts under the thin section are all unidentifiable, although rare
fossils may be bivalve shells. Porosity is variable (\textless1-10\%).

The consultation of the lithotheque hosted at UCA allowed us to identify
a subset of cherts with similarities to those from Type 7 from the
Middle Subbetic region (Betic Systems, south Spain,
Figure~\ref{fig-map}). These cherts are massive, sub-translucent,
peloidal/oolitic packstones. Allochems include very frequent peloids and
ooids. The peloids are densely arranged in the samples, more or less
visible depending on the alterations to the surface or the thickness.
The ooids are poorly sorted, with oval or round shapes, and replaced by
quartz. Uncommonly the ooids show a preserved nucleus. The presence of
bioclasts varies between samples, ranging between common to rare, and
when present/identifiable include sponge spicules (common), foraminifera
(rare to uncommon), and echinoderm spines (rare). Similar cherts have
also been identified and described in previous works, describing grey
cherts with peloidal and oolitic facies and occasional foraminifera
(Rodríguez, Rodríguez, and Pelegrin 2011). Despite the non-existence of
thin sections for comparison, the macroscopic similarities between the
cherts allow us to interpret Type 7 as coming from the Upper Jurassic
formations of the Betic Systems, located in the south and southeastern
Iberian Peninsula, at least 250 km (as the crow flies) east from Vale
Boi (Figure~\ref{fig-map}).

Type 8 (Figure~\ref{fig-distance-types}, e-f) corresponds to a massive
micro-cryptocrystalline quartz from a marine environment and
shelf/platform depositional environment. It represents
\textasciitilde2\% of the overall Terrace cherts and \textasciitilde5\%
of the chert in the Shelter. Identified and well-preserved fossils are
foraminifera from the Pfenderenidae family
(Figure~\ref{fig-distance-types} e, f2), which are larger benthic
foraminifera from marine environments (BouDagher-Fadel 2008). The
association of these foraminifera indicates this chert formed from
sediments deposited possibly between the Jurassic to the Cretaceous.
Macroscopically, this type is characterized by grey and white colours
(Figure~\ref{fig-distance-types} e-f1). The translucency ranges from
sub-translucent to opaque. Whenever present, the cortex is rounded and
thin (Table~\ref{tbl-cortex_tab}). Petrographically, this type is a
packstone (mudstone texture close to the edges) with a massive
microstructure. It is composed of 97\% micro-cryptocrystalline quartz,
macrocrystalline quartz, and fibrous chalcedony mostly replacing
fossils. Other minerals include uncommon iron oxides and the porosity is
of vug type (3\%). Fossil content is very frequent
(Figure~\ref{fig-distance-types} f2-f3), albeit mostly unidentifiable
and poorly preserved. Identified fossils are rare sponge spicules, rare
ostracods with some degree of preservation, common small, unidentified
foraminifera (100 μm) of differing preservation degrees, and as
mentioned above, foraminifera from the Pfenderenidae family (500 μm).
Despite the macroscopic similarities between type 8 and cherts from the
Middle Subbetic region chert formations (Figure~\ref{fig-map}), the
geological reference samples showed no macroscopically visible
foraminifera. However, our results indicate that despite no foraminifera
being visible macroscopically, they are present in thin section. The
fact that no foraminifera were identified in the small number of
analysed samples from the UCA lithotheque with similarities to Type 8
(n=3) may simply reflect their uncommon visibility, possibly due to
patination. Another possible source for Type 8 cherts may be the Upper
Cretaceous wackestone/packstone cherts with foraminifers from the Lisbon
region (Jordão and Pimentel 2022). However, these studies do not specify
the existence of Pfenderenidae foraminifers. Given the lack of this
specific fossil in the reference samples from the Betic chert
\textasciitilde250 km east and southeast, and the central Portugal
cherts 215-250 km north, the probable source of Type 8 remains unknown.

Type 9 (Figure~\ref{fig-distance-types} g) corresponds to a banded
micro-cryptocrystalline quartz/micrite peloidal packstone. It
corresponds to less than \textasciitilde2\% of the overall cherts in the
Terrace and Shelter areas. It is opaque, has a heterogeneous structure
and is characterized by a variable colour distribution: it shows a
horizontal, finely laminated pattern, with grey dark bands intercalated
with light grey bands (Figure~\ref{fig-distance-types} g1). Whenever
present, the cortex is rounded and is generally thin
(Table~\ref{tbl-cortex_tab}). In specific samples, the cortex is present
on two parallel planes of the sample (parallel to the laminations),
which indicates the chert was originally available in bedded layers.
Petrographically this chert is a banded packstone (banding of
sedimentary origin; Figure~\ref{fig-distance-types} g2), composed mostly
of micro-cryptocrystalline quartz (48\%), micrite (40\%), dolomite
(10\%), and accessory fibrous chalcedony and mica (muscovite). The
allochems are peloids, unidentifiable bioclasts, iron oxides, and opaque
minerals. Porosity is low (1\%). Although its origin is unknown, when
visiting the UCA lithotheque, we identified a small type of banded black
cherts without noticeable fossils or inclusions. Despite the lack of
comparative petrographic data, the macroscopic resemblances raise the
possibility that Type 9 belongs to Betic chert formations in southern
Spain, 250 km east (as the crow flies) from the site
(Figure~\ref{fig-map}).

Type 10 (Figure~\ref{fig-distance-types} h) is an oolitic packstone,
massive micro-cryptocrystalline micrite/quartz from a marine environment
and possibly a high-energy, shallow depositional environment, due to the
type, sorting, and preservation conditions of the ooids. It represents
\textasciitilde1\% of the overall Terrace cherts and less than 0.1\% of
the chert in the Shelter. Its heterogeneous structure has two types of
fabrics with different macroscopic characteristics: brown (opaque) and
black (sub-translucent) fabric (Figure~\ref{fig-distance-types} h1). The
sample chosen for the thin section is composed of 35\%
micro-cryptocrystalline quartz, 53\% micrite, 9\% macrocrystalline
quartz, and accessory fibrous chalcedony (filling fractures) and mica;
the percentages of quartz and micrite might be related to the fabrics
present in the artefacts. Allochems are common iron oxides and very
frequent ooids. Bioclasts are rare, poorly preserved, and replaced by
quartz. The porosity is intraparticle and vug (10\%). The ooids are
distributed homogeneously across the sample (although macroscopically
visible in the brown fabric) and are highly abundant and concentrated.
They are poorly sorted, varying between 500 and 20 μm
(Figure~\ref{fig-distance-types} h2), show round-to-elliptical shapes
and their preservation is variable: some ooids show a poorly preserved
micritic structure, while others show concentric laminae structures
around a round nucleus (Flügel 2010). This chert shows no similarities
to any consulted reference material. Despite the existence of oolitic
limestones in the Algarve region correspondent to Middle and Upper
Jurassic formations and often described as whiteish (Rocha et al.,
1979), the presence of oolitic chert in the region is not recognized in
geological and archaeological literature (Cardoso, Andrade, and Martins
2018; Rocha et al. 1979; Verissimo 2004) and the visit of outcrops
mentioned in the geological cartography did not allow their
identification.

Type 11 (Figure~\ref{fig-small-types}) is a banded/laminated mudstone,
characterized by the presence of a conchoidal fracture (within the
nodule) and a non-conchoidal, laminar fracture (at the edges of the
nodule). This chert type is grey, opaque, and without identifiable
allochems. This chert shows a limited presence at the archaeological
site as previously mentioned, restricted to the Solutrean occupations,
and corresponding to less than 0.1\% of the Terrace cherts and
\textasciitilde2\% of Shelter cherts.

\begin{figure}

\centering{

\includegraphics{figures/within-and-beyond/small_grid.png}

}

\caption{\label{fig-small-types}Macroscopic view with a stereomicroscope
of non-local Type 11 and TL cherts. Type 11 -- detail of banded patterns
composed of parallel different fabrics; TL01 -- detail of the jasper and
fractures filled with chalcedony; TL02 -- detail of the chalcedony with
unidentified orange inclusions; TL03 -- detail of two round inclusions,
possibly iron oxides; TL04 - detail of massive chert with fractures
filled with chalcedony; TL05 -- detail of the banded pattern within the
grey mudstone texture; TL06 - detail of the small, finely sorted ooids;
TL07 -- detail of the dendritic laminations, possibly composed of iron
oxides; TL08 -- detail of the mudstone structure, without the presence
of allochems; TL09a/b -- detail of the mossy, orange and red inclusions
within the chalcedony; TL10 -- detail of the speckling and laminated
patterns within the mudstone structure; TL11 -- detail of the banding
within the mudstone structure; TL12 -- detail of the poorly preserved
ooids, only visible under the stereomicroscope; TL13 -- detail of the
bands and laminations within the mudstone texture fabric; TL14 -- detail
of an unidentified benthic foraminifera; TL15 -- detail of the interface
between the massive brown fabric and splotches; TL16 - detail of the of
the massive structure with unidentifiable inclusions.}

\end{figure}%

Finally, the TL types (Figure~\ref{fig-small-types}) are characterized
by several samples without macroscopic correspondence to the previously
identified types or to the local cherts and chalcedony from the Algarve
region previously described by Belmiro et al. (2023). These are composed
mostly of blanks or retouched tools, rarely with cortex
(Table~\ref{tbl-cortex_tab}), and rarely the same TL type is found in
different archaeological layers. A detailed description of these types
can be found in the Supplementary Information (Online Resource 2). In
both areas, the TL cherts compose less than 1\% of the cherts.

Apart from the types common throughout the stratigraphy of the Terrace
and Shelter, Type 11 and samples from the TL type also show no
identifiable source. The exceptions are TL01, TL10-11 and TL15-16.

Jasper is known in the South Portuguese Zone area (Oliveira 1984), from
southern/central Portugal to Spain, north of the Algarve (from Aljustrel
to Beja; Figure~\ref{fig-map}), suggesting that the source of TL01 may
be from south/central Portugal at a minimum distance of 100 km north and
northeast from Vale Boi (as the crow flies).

Similarly, despite being different from the most frequent Cretaceous
nodules, the macroscopic similarities between TL10 and TL11 and specific
nodules from central Portugal in the Lusolit lithotheque may indicate
these cherts are from the Estremadura/Lisbon area (Figure~\ref{fig-map}
a). This difference in appearance may be due to the already-mentioned
variability of the nodule's macroscopic characteristics.

Finally, TL15 and TL16 show macroscopic similarities to radiolarite
reference samples located at the Cadiz lithotheque, which may indicate
the source of these cherts is from the Middle Subbetic region (Domínguez
Bella 2010; Domínguez-Bella 2006).

\subsection{Stratigraphical
distribution}\label{stratigraphical-distribution}

\subsubsection{Terrace area}\label{terrace-area}

Most chert types described above are consistently present across the
Terrace stratigraphy (except for Type 11 which appears only in level 4
and in a small number of artefacts; n=2). However, being ubiquitous, we
identified a difference in the percentages of chert types throughout the
stratigraphy (Figure~\ref{fig-type-per-level}).

In the Gravettian occupations (levels 6 and 7) and at the start of the
Proto-Solutrean (level 5) local cherts make up \textasciitilde45-60\%,
while the combined non-local cherts make up \textasciitilde65-40\%.
Level 6 shows the highest percentages of all non-local chert use, with
specific types like Type 6 representing \textasciitilde30\%, although
when considering the sum of weight, the percentages reduce, and
non-local chert represents \textasciitilde50\% of total chert. The
percentages of local chert increase in the following occupations. In the
Proto-Solutrean (levels 4E and 4E/5) and Solutrean (levels 4, 4C, and
4D) local cherts make up more than \textasciitilde70\% of the sample and
80\% when considering the sum of weight, while non-local cherts decrease
significantly and represent percentages lower than 30\%. Despite the
increase of local cherts during the Proto-Solutrean and Solutrean, there
are still significant differences between the local types present in the
two technocomplexes. During the top Proto-Solutrean occupations the
grey/white cherts (Type 4) are the most present chert with percentages
between 35-20\%, while in the Solutrean levels, the yellow/brown cherts
(Type 2) are present in the highest percentage, at \textasciitilde30\%.

\begin{figure}

\centering{

\includegraphics{article2_files/figure-pdf/fig-type-per-level-1.pdf}

}

\caption{\label{fig-type-per-level}Terrace area chalcedony and chert
type frequency (a) and weight sum (b) per level, without unidentifiable
samples. Only percentages superior to the 5\% threshold are shown in the
figure. Type 1 to 5 (yellow and red colours) represent local chalcedony
and cherts. Type 6, 7 and 9 represent non-local cherts with identified
probable provenience (brown colours). Type 8, 10 and 11 represent
non-local cherts with non-identified provenience (blue and green
colours).}

\end{figure}%

When plotting the weight, it is apparent that most of the samples are
located under the 10 gr mark, with median and bottom quartiles often
falling under the 5 gr line (Figure~\ref{fig-weight}). Local types, such
as Type 2, Type 4, and Type 5, although mostly represented by artefacts
with a medium mass (\textasciitilde5-10 gr), show the heaviest and
highest number of heavy artefacts. In comparison, other types, although
showing some variation in the weight of artefacts, show smaller
artefacts and their means and medians are frequently closer to the
bottom quartile. As seen in (Figure~\ref{fig-weight}), non-local types
show larger weights from level 5 to level 7.

\begin{figure}

\centering{

\includegraphics{article2_files/figure-pdf/fig-weight-1.pdf}

}

\caption{\label{fig-weight}Terrace boxplots of chert type weight (gr)
per level. Diamond shape within the boxplots represents weight mean.
Type 1 to 5 (yellow and red colours) represent local chalcedony and
cherts. Type 6, 7 and 9 represent non-local cherts with identified
probable provenience (brown colours). Type 8, 10 and 11 represent
non-local cherts with non-identified provenience (blue and green
colours).}

\end{figure}%

\subsubsection{Shelter area}\label{shelter-area}

The same raw materials found in the Terrace area are also found in the
Solutrean occupation of the Shelter, albeit with some differences
(Figure~\ref{fig-analysis_shelter}). The local cherts make up between
\textasciitilde65-75\% (considering both counts and weight) of the
sample and are mainly composed of the brown/yellow cherts (Type 2) that
frequently characterize the Lower Jurassic outcrops of the Algarve
(Figure~\ref{fig-analysis_shelter} a-b). The presence of cherts not
congruent with the local sources is, like level 4 in the Terrace, close
to 25\%, although dominated mostly by Type 6 (\textasciitilde15\%).
Similarly to the Solutrean occupations of the Terrace, although the
artefacts are generally small (lower weights), the local types show a
higher variability in mass, with several outliers and heavy samples when
compared to the non-local types (Figure~\ref{fig-analysis_shelter} c).

\begin{figure}

\centering{

\includegraphics{article2_files/figure-pdf/fig-analysis_shelter-1.pdf}

}

\caption{\label{fig-analysis_shelter}a-b) Shelter area chalcedony and
chert type frequency (a) and weight sum (b) per level, without
unidentifiable samples. Only percentages superior to the 5\% threshold
are shown in the figure. c): Shelter boxplots of chert type weight (gr)
per level. Type 1 to 5 (yellow and red colours) represent local
chalcedony and cherts. Type 6, 7 and 9 represent non-local cherts with
identified probable provenience (brown colours). Type 8, 10 and 11
represent non-local cherts with non-identified provenience (blue and
green colours).}

\end{figure}%

\section{Discussion}\label{discussion}

The macroscopic and petrographic analysis of chert and chalcedony lithic
materials from the UP levels of the Terrace and Shelter areas at Vale
Boi yielded essential insights into the procurement of raw materials and
the mobility patterns of hunter-gatherers in southwestern Iberia. These
materials serve as valuable indicators for identifying changes in
mobility, as they reveal patterns of both local and non-local
procurement and have allowed us to answer questions about landscape
exploitation and territory movement through the presence (or absence) of
local, regional, and long-distance cherts.

The patterns of high percentages of local raw materials in all
assemblages indicate a heavy reliance on nearby resources. The local
chalcedony and chert indicate that the groups were sourcing a
significant portion of their raw materials within approximately a 20 km
radius of the site. The direct procurement of these materials is
plausible, given that known chalcedony primary sources are situated to
the west, about 10 km from Vale Boi (or possibly in the vicinity of the
site), and chert sources are found to the east, within the inland Lower
Jurassic outcrops, roughly 8 km away (Figure~\ref{fig-map} b). This
hypothesis is supported by the potential existence of UP occupations
near these outcrops, which might reflect quarrying and chert catchment
activities during prehistory (Zambujo 1998). Evidence of short-distance
procurement at Vale Boi includes the presence of heavy nodules
(Figure~\ref{fig-weight}) with little indication of knapping and large
chunks of chert from local varieties, often with parent rock still
attached.

It is also possible that local cherts were procured from farther
distances, perhaps as part of embedded activities (Pereira, Bicho, et
al. 2016). The chert sources from the Lower and Upper Jurassic periods
in Western Algarve, located along the beach and cliff areas, can be
found between 15-18 km southwest of Vale Boi (Figure~\ref{fig-map} b).
While our current data is not able to pinpoint which specific outcrops
were exploited based on macroscopic and petrographic analysis, the
identification of persistent coastal adaptions at Vale Boi (Nuno Bicho
and Esteves 2022) and exploitation of marine resources such as shells at
the site indicates that these groups visited the coast and used the
available coastal resources throughout all of the UP (Belmiro et al.
2021; Manne and Bicho 2011). This suggests that raw materials could have
been conveniently acquired in coastal vicinities without significant
detours (Binford 1979; Surovell 2009).

The limitations of macroscopic and petrographic methods in discerning
between outcrops due to the chert nodule similarities have also impaired
the specific distinction between Lower Jurassic and Upper Jurassic
sources in western Algarve (within approximately a 20 km radius of the
site) and Middle Jurassic cherts located in eastern Algarve
(approximately 80 km to the east; Figure~\ref{fig-map} c). As such, the
presence of these regional resources at Vale Boi is uncertain.

A similar issue was noted by Surovell (2009), who described a scenario
at the Late Pleistocene archaeological site of Krmpotich (USA), where a
specific type of chert found at the site could be sourced from both
local and long-distance outcrops. The author suggested that some of
these cherts might have been imported from distant sources or previous
campsites.

Consequently, we cannot rule out the possibility that some of the cherts
we have identified as being sourced within a 20 km radius of Vale Boi
may actually be the result of regional procurement and mobility,
potentially from areas approximately 80-100 km away, where the Middle
Jurassic cherts are located. Given the similarity of the raw materials
and the absence of geochemical studies which have shown to be key to
differentiating between cherts belonging to the same geological
formations or cherts formed under similar conditions (Gómez De Soler et
al. 2023), and considering the studies and model mentioned above, most
of these cherts likely originate from local sources, while a minor
portion may be attributable to regional movements.

In contrast to the Middle Jurassic cherts, the regional Upper Jurassic
cherts located \textasciitilde90 km east of Vale Boi
(Figure~\ref{fig-map} c) have easily discernible characteristics.
However, they have not been detected in the assemblages. This absence
could be attributed to the limited use and transport of these nodules
from their sources to Vale Boi, potentially due to the small size of the
nodules (often between 3-5 cm), the difficulty of detaching them from
the parent rock, or challenges related to their accessibility and
visibility, given that the known outcrops are situated inland,
approximately 10 km from the present coastline. This may corroborate the
previously observed interest of Pleistocene hunter-gatherers in coastal
environments in Iberia (Schmidt et al. 2012), which would have been
characterized by rich marine productivity due to upwelling especially in
the Atlantic coast (Nuno Bicho and Esteves 2022; N. Bicho and Haws
2008), as well as the aforementioned adaptations and use of coastal
resources of hunter-gatherers from Vale Boi (Nuno Bicho and Esteves
2022; Manne and Bicho 2011).

A similar case is observed for the jaspers located in the SPZ (Oliveira
1984), located at a minimum of 100 km distance from Vale Boi, as the
crow flies. Despite being a source located at closer distances when
compared to the cherts of central Portugal or from southwestern Spain,
the jasper occurs in trace amounts (n=3) and only in the Gravettian
layers (layer 6). Occurring north and northeast, the limited amount of
jasper in the assemblage may either represent a limited use of these
resources in detriment of the cherts from the Lisbon and Rio Maior area,
or suggest that they were used and knapped elsewhere, but not
transported, either by the same groups or through exchanges by different
groups.

So far, we have discussed the presence of cherts whose outcrops are in
relative proximity to Vale Boi and show the full usage of local chert
variability (albeit with uncertainty regarding the Upper Jurassic cherts
of western Algarve due to their macroscopic similarities) and the
selective usage of regional cherts within the Algarve. As previously
mentioned, however, it is the combined identification of local and
long-distance cherts that allows us to have a better idea of the total
exploited territory, as well as the possible existence of exchanges and
social networks between groups occupying different territories.

Although a wide variety of chert types incongruent with the regional
reference collection were identified throughout the stratigraphy and
classified as non-local, we were only able to ascertain the possible
sources of three types: Type 6, Type 7 and Type 9. The long-distance
cherts with identified sources come from distances of over
\textasciitilde200 km and in two main directions: 1) from the north,
where Cretaceous cherts from central Portugal are found
(Figure~\ref{fig-map} a); and 2) from the east, where peloidal/oolitic
chert from western Andalusia is present (Figure~\ref{fig-map}), and
possibly jasper, transported by the Guadiana river at the current border
between Portugal and Spain.

In fact, Vale Boi's proximity to a coastal setting, with access to
beaches, as well as rivers such as the Guadiana River, where raw
material nodules could be transported from far-away distances, could
explain the presence of long-distance cherts. In this case, the presence
of cherts with sources from over 200 km would be explained by their
procurement in secondary settings and a possible local movement range.
However, the scarcity of pebble cortex and the predominance of rounded
cortex suggest that these cherts were likely not collected from beaches
or riverbeds far from the outcrops but were possibly gathered closer to
the source areas (Kuhn 2004). As such, the presence of these
long-distance cherts is better explained by the extended movement of
human groups in the south-north and west-east axes.

However, it is relevant to ask why human groups occupying Vale Boi were
obtaining a considerable amount of their chert from such long distances
and such specific directions, either directly through procurement or
through exchanges with other groups. Although the source of the other
non-local cherts is still unknown, making the directions and distances
to obtain them also unknown, the identified axes of long-distance
movement to the north and east may be the result of the influence of the
topography on mobility patterns. Indeed, formal models that seek to
explain hunter-gatherer mobility have highlighted the significance of
factors such as topography on the distances travelled by groups or
individuals (Brantingham 2006).

Similarly, friction of terrain cartography for Iberia has shown that
terrain can produce different costs of travelling on foot, which differ
from those directly observed on current-day maps and direct distances
(Díaz del Río 2020). In this context, distance is less related to
measurements and more related to the time it costs to go from point A to
point B (Zilhão 2021). For example, Díaz del Río (2020) highlights the
existence of low-cost pathways through plateaus and the connectivity
between the river valleys of central and south Iberia (Guadalquivir,
Guadiana and Tagus).

Using the same friction of terrain cartography, Zilhão (2021) also
identifies the possible role of terrain in hunter-gatherer mobility in
the Upper Paleolithic. For example, the author highlights that the
Parpalló bifacial point is found across central/south Spain and Portugal
during the Upper Solutrean, and thus is considered the index fossil for
this period in the region. From westernmost central Portugal to
easternmost central Spain (west-east axis across Iberia), there is a
distance of \textasciitilde800-900 km as the crow flies. However, the
Parpalló point is not found in northern Spain (Catalunya), only 200 km
north of the easternmost site where the index fossil was identified. The
author suggests that this separation may be explained by the friction of
terrain, which divides Iberia into two halves, creating a quicker
mobility pathway between central/south Spain and Portugal in opposition
to the higher-cost time necessary to travel to northern Spain.

As such, the observed long-distance transfer of cherts in the north/east
directions by hunter-gatherers at Vale Boi might reflect this central
and southern Iberia territory dynamic and the exploitation of the
natural corridors for movement, explaining the substantial use of
long-distance chert. In this case, the presence of a smaller subset of
non-local cherts of unknown origin (TL) but not knapped at the site
could represent different forms of long-distance trade of finished goods
and blanks, or interactions with various social networks, potentially
even indicating individual networks and mobility strategies (sensu
Gamble 1999).

Another key idea about these mobility patterns is not only their
distance and direction but also their continuity throughout the UP. The
detection of the same types of non-local raw materials throughout the
stratigraphy at Vale Boi points to a persistent use of long-distance
chert sources. This continuity cannot only be explained by the natural
topographic connectivity between central and south Iberia but can also
be attributed to a tradition of shared knowledge and culture regarding
the locations of raw materials and other resources within the landscape,
or the preservation of social networks over time.

For example, Cascalheira et al. (2017) acknowledge the adoption of
specific behavioural strategies by Vale Boi hunter-gatherers in response
to the regional ecological conditions, which were sustained throughout
the UP at the site. Practices such as grease rendering or the functional
specialization of raw materials (e.g., the use of quartz and greywacke
for bone breakage activities or quartz for grease rendering) are seen as
deliberate actions with significant cultural implications for future
generations, influencing the adaptive choices made over time.
Accordingly, chert raw material acquisition and management appear to be
integral to these behavioural strategies, established from the site's
earliest occupations and perpetuated throughout the UP.

However, despite the consistent presence of local and long-distance
chert throughout the stratigraphy, we identified trends of changing
frequencies in the different occupations, possibly also providing
different interpretations about mobility and social behaviour through
time. For example, the significant presence of non-local cherts during
the Gravettian occupations at Vale Boi suggests extensive communication
and mobility among the groups occupying the site. As previously
established, however, while it would be expected for Solutrean groups to
have high mobility and extended social networks, the opposite would be
expected for the Gravettian occupations. As such, our results,
especially those from the Gravettian occupations, diverge from the
anticipated patterns.

Specifically, technological analyses of lithic assemblages have unveiled
notable differences between Gravettian assemblages from the Atlantic and
Mediterranean coastal regions, pointing to the existence of two distinct
regional Gravettian traditions in southern Iberia: the Vicentine
Gravettian at Vale Boi, and the Mediterranean Gravettian identified in
UP sites in southern Spain (Marreiros et al. 2015; Marreiros and Bicho
2013). These traditions are thought to reflect adaptations to the unique
local and regional ecological settings (Bradtmöller et al. 2016) and
evolving environmental conditions (Marreiros and Bicho 2013),
potentially fostering distinct ethnographic divisions among
hunter-gatherer communities (Marreiros et al. 2015). Consequently, it
would be expected for territorial or cultural demarcations to influence
material usage patterns, with a shift towards a greater reliance on
local resources and a diminished use of non-local raw materials during
the Gravettian at Vale Boi.

However, the observed prevalence of non-local chert materials could
reflect not merely long-distance movements within a single cultural
framework but rather the result of interactions and exchanges between
different cultural groups, thus explaining the existence of distinct
regional technological characteristics.

This perspective aligns with Brantingham's (2006) discussion on social
exchange, which posits that exchanges likely occurred between groups
occupying distinct territories, who would meet regularly to trade raw
materials. This model helps explain the presence of materials from
distant sources at archaeological sites, illustrating the nuanced and
interconnected nature of prehistoric social landscapes (Whallon 2006).

In this sense, interpreting the ubiquitous presence of long-distance
chert at Vale Boi as a result of exchange between groups and social
networks informs us that even during chronological periods where
ethnographic boundaries have been suggested, there is communication and
exchange between groups. This further cements the idea of
interconnectedness across the Iberian territory during the UP, expanding
the extensive line of contacts and networks identified across north and
central Iberia (Aubry et al. 2004, 2016; Corchón Rodríguez, Martínez,
and Tarriño 2016; Hermida, Rellán, and Rodríguez 2016; Marreiros et al.
2016; Ortega 2003; Sánchez de la Torre et al. 2023) to southwesternmost
Iberia.

However, although social networks may explain the presence of
long-distance cherts at Vale Boi even during periods of demarked
regional boundaries, the high/low presence of long-distance raw
materials at the site may not be the direct result of a higher reliance
on these networks, or their expansion or shrinkage.

An alternative interpretation of the varying frequencies of non-local
cherts during different periods at Vale Boi can be explained by the
``Mean per Capita Occupation Span'' model proposed by Surovell (2009),
and the ``Distance/Frequency Index (DFI)'' outlined by Grove et al.
(2023). These models hypothesize that shorter-term occupations are
likely to show a higher proportion of non-local artefacts, whereas with
longer-term settlements, the prevalence of local artefacts increases.
This pattern may be attributed to environmental influences: challenging
conditions such as fragmented habitats and cooler temperatures limit
mobility, making long-term settlement in areas rich in local resources a
preferable, lower-risk option (Surovell 2009).

This pattern is mirrored in the archaeological record of Vale Boi,
where, aside from the earlier Gravettian of levels 6 and 7, subsequent
occupations seem to adopt a logistical-settlement strategy, positioning
Vale Boi as a long-term basecamp (Cascalheira et al. 2017). In this
context, the Gravettian phase at Vale Boi could reflect a period of
enhanced mobility and brief occupations, possibly interspersed with
reoccupations, indicative of a significant import of non-local raw
materials. This scenario parallels observations from the Late Gravettian
in the French Massif central, where hunter-gatherer groups are known to
have transported entire or pre-formed blocks of raw material over
considerable distances (\textgreater{} 200 km) for short-term,
hunting-related site occupations (Delvigne et al. 2019). The authors
also note that in the following occupations, during the Badegoulian and
Magdalenian, occupation duration increases along with a significant
increase of local and semi-local raw materials.

Similarly, the lower percentages of non-local cherts characterizing the
phases following the Gravettian at Vale Boi could reflect a period of
increased site occupation. In this context, and following the model by
Surovell (2009), the shift toward colder and drier climatic conditions
marked by Heinrich Event 2 (HE2) during the Proto-Solutrean, followed by
the Last Glacial Maximum in the Solutrean period (Belmiro et al. 2021;
Cascalheira and Bicho 2013; González-Sampériz et al. 2010; Sanchez Goñi
and Harrison 2010; Schmidt et al. 2012), could have resulted in a more
challenging environment with sparser vegetation (Cascalheira and Bicho
2013; González-Sampériz et al. 2010). These changes likely constrained
the mobility of hunter-gatherers, fostering longer-term settlements
within the resource-rich niche of southwestern Portugal (Cascalheira et
al. 2017; Schmidt et al. 2012). Consequently, as settlements extended in
duration, the reliance on non-local raw materials, whether sourced
directly or acquired through trade, diminished in favour of increasingly
dominant local raw materials.

\section{Conclusions}\label{conclusions}

Our findings build upon earlier studies indicating that the majority of
chert utilized at Vale Boi originates from local Lower Jurassic sources
within a 20 km radius of the site (Pereira, Bicho, et al. 2016), with
non-local materials constituting a relevant portion of the chert used.
The hypothesis of long-distance procurement of Cretaceous cherts from
central Portugal, situated north of Vale Boi, is corroborated by visual
and petrographic evidence (Pereira, Bicho, et al. 2016; Nuno Bicho et
al. 2013), and this pattern persists throughout the UP. Furthermore, the
identification of cherts bearing macroscopic resemblance to the
peloidal/oolitic Upper Jurassic cherts from the Betic System in southern
Spain confirms a west-east mobility axis, previously inferred from
techno-typological studies, also spanning the entirety of the UP period.

These findings reveal a landscape of hunter-gatherers engaged in complex
practices of resource management, and possibly risk management. The
presence of allochthonous raw materials, albeit in small quantities
during the Proto-Solutrean and Solutrean phases, underscores that the
inhabitants of Vale Boi---situated on the periphery of Iberia and
Europe, bordered by the sea to the south and west, and confined to the
patchy refugia created by the worsening climatic conditions of the HE2
and following Glacial Maximum (Schmidt et al. 2012)---were certainly not
isolated. Instead, they were part of an extensive social network, a
pattern of interconnectivity that resonates with findings across western
Europe (Hermida, Rellán, and Rodríguez 2016; Corchón Rodríguez,
Martínez, and Tarriño 2016; Ortega 2003; Aubry et al. 2016, 2004;
Sánchez de la Torre et al. 2023).

In southern Portugal, this interconnection among UP hunter-gatherers
from diverse and distant regions of Iberia is similarly observed. This
was already expected for the Proto-Solutrean and Solutrean occupations,
but less so for the Gravettian where social boundaries were expected
based on previously identified technological facies between regions
(Marreiros et al. 2015). Instead, our results show a high presence of
long-distance raw materials during the Gravettian occupations, which may
indicate that social networks were established from the first UP
occupations in south Portugal and maintained through time.

The outcomes of this research pave the way for continued exploration of
these assemblages through diverse methodologies. Future directions
include employing a geochemical approach to both the studied assemblages
and the geological reference collection, aiming to enrich discussions on
mobility throughout the UP at Vale Boi. This approach is anticipated to
provide diagnostic elements to distinguish between cherts from Lower and
Middle Jurassic units, thereby offering a comprehensive view of regional
resource utilization and mobility patterns.

Additionally, experimental replication and the creation of a reference
collection modified through heat treatment could enhance the
identification of raw material types and currently indistinguishable
samples, as noted in prior research (Pereira, Bicho, et al. 2016).
Lastly, we intend to integrate technological data from current and
ongoing datasets to delve deeper into the usage and management of raw
materials (both local and non-local) over time, their association with
different site uses, and to contextualize these practices within the
broader framework of human adaptations to changing environmental
conditions throughout the UP.

\section*{Supplementary Information}\label{supplementary-information}
\addcontentsline{toc}{section}{Supplementary Information}

\markright{Supplementary Information}

\textbf{Online Resource 1}: Data dictionaries for macroscopic (group and
individual) and petrographic analyses.

\textbf{Online Resource 2}: Terrace and shelter lithotypes complete
macroscopic and petrographic description.

\textbf{Online resource 3}: Summarised type macroscopic and petrographic
features and epoch/stage identification.

\bookmarksetup{startatroot}

\chapter{geochemistry}\label{geochemistry}

\section{X-ray diffraction (XRD)}\label{x-ray-diffraction-xrd}

Some text. Some other text. A bit more of text.

\section{Scanning electron microscopy and energy dispersive X-ray
spectroscopy
(SEM-EDS)}\label{scanning-electron-microscopy-and-energy-dispersive-x-ray-spectroscopy-sem-eds}

\section{Portable X-ray fluorescence
(pXRF)}\label{portable-x-ray-fluorescence-pxrf}

\begin{figure}

\centering{

\includegraphics{geochemistry_files/figure-pdf/fig-lod-1.pdf}

}

\caption{\label{fig-lod-1}A barplot.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{geochemistry_files/figure-pdf/fig-lod-2.pdf}

}

\caption{\label{fig-lod-2}A barplot.}

\end{figure}%

\begin{verbatim}

Call:
PCA(X = df_scaled_alg, graph = FALSE)

Eigenvalues
                       Dim.1   Dim.2   Dim.3   Dim.4   Dim.5   Dim.6   Dim.7
Variance               2.419   1.842   1.009   0.800   0.663   0.151   0.116
% of var.             34.558  26.319  14.408  11.424   9.477   2.154   1.660
Cumulative % of var.  34.558  60.877  75.284  86.708  96.186  98.340 100.000

Individuals (the 10 first)
          Dist    Dim.1    ctr   cos2    Dim.2    ctr   cos2    Dim.3    ctr
LJW   |  5.517 |  2.737  4.995  0.246 | -1.311  1.504  0.056 |  2.627 11.035
LJW.1 |  3.781 |  3.129  6.527  0.685 | -1.509  1.995  0.159 | -0.957  1.465
LJW.2 |  5.036 |  2.464  4.047  0.239 |  2.696  6.365  0.287 |  2.566 10.531
LJW.3 |  1.616 |  0.869  0.503  0.289 | -1.289  1.455  0.636 | -0.092  0.013
LJW.4 |  6.734 |  3.205  6.849  0.227 |  4.295 16.149  0.407 |  3.622 20.978
LJW.5 |  3.092 | -2.104  2.952  0.463 |  2.084  3.804  0.455 | -0.332  0.176
LJW.6 |  2.173 |  1.088  0.789  0.251 | -1.751  2.683  0.649 | -0.387  0.240
LJW.7 |  1.942 |  1.247  1.036  0.412 | -1.282  1.440  0.436 | -0.305  0.148
LJW.8 |  3.021 |  2.064  2.840  0.467 | -1.639  2.352  0.294 | -0.995  1.582
LJW.9 |  6.299 |  4.211 11.824  0.447 |  3.056  8.174  0.235 | -1.359  2.953
        cos2  
LJW    0.227 |
LJW.1  0.064 |
LJW.2  0.260 |
LJW.3  0.003 |
LJW.4  0.289 |
LJW.5  0.012 |
LJW.6  0.032 |
LJW.7  0.025 |
LJW.8  0.108 |
LJW.9  0.047 |

Variables
         Dim.1    ctr   cos2    Dim.2    ctr   cos2    Dim.3    ctr   cos2  
P2O5  |  0.805 26.793  0.648 |  0.490 13.038  0.240 | -0.148  2.184  0.022 |
S     |  0.659 17.966  0.435 |  0.491 13.067  0.241 | -0.154  2.358  0.024 |
Cl    |  0.360  5.358  0.130 |  0.474 12.190  0.225 |  0.611 37.034  0.374 |
K2O   |  0.667 18.400  0.445 | -0.675 24.757  0.456 | -0.141  1.971  0.020 |
CaO   |  0.081  0.270  0.007 |  0.596 19.287  0.355 | -0.379 14.249  0.144 |
TiO2  |  0.730 22.056  0.534 | -0.530 15.270  0.281 | -0.223  4.925  0.050 |
Fe2O3 |  0.471  9.157  0.222 | -0.210  2.391  0.044 |  0.613 37.279  0.376 |
\end{verbatim}

\includegraphics{geochemistry_files/figure-pdf/PCA-1.pdf}

\begin{verbatim}

Call:
PCA(X = df_scaled_nl, graph = FALSE)

Eigenvalues
                       Dim.1   Dim.2   Dim.3   Dim.4   Dim.5   Dim.6
Variance               2.933   1.713   0.672   0.354   0.249   0.079
% of var.             48.885  28.546  11.197   5.901   4.157   1.314
Cumulative % of var.  48.885  77.431  88.628  94.529  98.686 100.000

Individuals (the 10 first)
          Dist    Dim.1    ctr   cos2    Dim.2    ctr   cos2    Dim.3    ctr
CPT   |  1.744 | -1.714  1.411  0.967 | -0.066  0.004  0.001 | -0.142  0.042
CPT.1 |  1.797 | -1.356  0.884  0.570 |  0.402  0.133  0.050 |  1.046  2.295
CPT.2 |  1.426 | -1.330  0.850  0.870 | -0.353  0.103  0.061 | -0.060  0.008
CPT.3 |  1.825 | -1.803  1.561  0.975 | -0.070  0.004  0.001 | -0.084  0.015
CPT.4 |  1.815 | -1.703  1.393  0.881 | -0.332  0.090  0.033 | -0.360  0.272
CPT.5 |  1.224 | -0.702  0.237  0.329 | -0.651  0.348  0.282 | -0.659  0.911
CPT.6 |  1.542 | -1.503  1.085  0.951 | -0.100  0.008  0.004 | -0.197  0.081
CPT.7 |  1.883 | -1.787  1.534  0.901 | -0.216  0.038  0.013 | -0.432  0.391
CPT.8 |  1.592 | -1.544  1.145  0.941 |  0.056  0.003  0.001 |  0.099  0.020
CPT.9 |  1.712 | -1.523  1.114  0.792 |  0.441  0.160  0.066 |  0.475  0.474
        cos2  
CPT    0.007 |
CPT.1  0.339 |
CPT.2  0.002 |
CPT.3  0.002 |
CPT.4  0.039 |
CPT.5  0.290 |
CPT.6  0.016 |
CPT.7  0.053 |
CPT.8  0.004 |
CPT.9  0.077 |

Variables
         Dim.1    ctr   cos2    Dim.2    ctr   cos2    Dim.3    ctr   cos2  
P2O5  |  0.668 15.194  0.446 |  0.591 20.375  0.349 | -0.119  2.094  0.014 |
K2O   |  0.836 23.802  0.698 | -0.463 12.504  0.214 | -0.073  0.791  0.005 |
TiO2  |  0.886 26.789  0.786 | -0.370  7.986  0.137 | -0.053  0.411  0.003 |
Fe2O3 |  0.844 24.304  0.713 | -0.237  3.278  0.056 |  0.110  1.797  0.012 |
Cu    |  0.411  5.761  0.169 |  0.632 23.299  0.399 |  0.637 60.406  0.406 |
S     |  0.349  4.150  0.122 |  0.747 32.557  0.558 | -0.481 34.501  0.232 |
\end{verbatim}

\includegraphics{geochemistry_files/figure-pdf/PCA-2.pdf}

\begin{verbatim}

Call:
PCA(X = df_scaled_comp, graph = FALSE)

Eigenvalues
                       Dim.1   Dim.2   Dim.3   Dim.4   Dim.5   Dim.6
Variance               2.729   1.753   0.712   0.456   0.247   0.103
% of var.             45.477  29.222  11.871   7.600   4.121   1.709
Cumulative % of var.  45.477  74.699  86.570  94.170  98.291 100.000

Individuals (the 10 first)
          Dist    Dim.1    ctr   cos2    Dim.2    ctr   cos2    Dim.3    ctr
CPT   |  1.872 | -1.795  1.639  0.919 |  0.369  0.108  0.039 | -0.111  0.024
CPT.1 |  1.907 | -1.389  0.982  0.531 |  0.687  0.374  0.130 |  0.936  1.710
CPT.2 |  1.356 | -1.275  0.827  0.884 | -0.107  0.009  0.006 |  0.041  0.003
CPT.3 |  2.000 | -1.935  1.907  0.936 |  0.426  0.144  0.045 | -0.116  0.026
CPT.4 |  1.850 | -1.773  1.600  0.918 |  0.089  0.006  0.002 | -0.328  0.209
CPT.5 |  1.308 | -0.359  0.066  0.075 | -0.852  0.574  0.424 | -0.776  1.174
CPT.6 |  1.570 | -1.502  1.149  0.915 |  0.237  0.045  0.023 | -0.150  0.044
CPT.7 |  1.991 | -1.882  1.802  0.894 |  0.248  0.049  0.016 | -0.399  0.310
CPT.8 |  1.674 | -1.552  1.225  0.860 |  0.353  0.099  0.044 |  0.044  0.004
CPT.9 |  1.986 | -1.618  1.333  0.664 |  0.979  0.759  0.243 |  0.475  0.440
        cos2  
CPT    0.003 |
CPT.1  0.241 |
CPT.2  0.001 |
CPT.3  0.003 |
CPT.4  0.031 |
CPT.5  0.352 |
CPT.6  0.009 |
CPT.7  0.040 |
CPT.8  0.001 |
CPT.9  0.057 |

Variables
         Dim.1    ctr   cos2    Dim.2    ctr   cos2    Dim.3    ctr   cos2  
P2O5  |  0.663 16.108  0.440 |  0.620 21.955  0.385 | -0.190  5.080  0.036 |
K2O   |  0.712 18.598  0.507 | -0.630 22.626  0.397 | -0.108  1.653  0.012 |
TiO2  |  0.800 23.482  0.641 | -0.491 13.763  0.241 | -0.048  0.321  0.002 |
Fe2O3 |  0.751 20.672  0.564 | -0.247  3.491  0.061 |  0.295 12.221  0.087 |
Cu    |  0.463  7.866  0.215 |  0.576 18.943  0.332 |  0.617 53.383  0.380 |
S     |  0.602 13.273  0.362 |  0.581 19.222  0.337 | -0.441 27.342  0.195 |
\end{verbatim}

\includegraphics{geochemistry_files/figure-pdf/PCA-3.pdf}

\begin{verbatim}

Call:
PCA(X = df_scaled_comp2, graph = FALSE)

Eigenvalues
                       Dim.1   Dim.2   Dim.3   Dim.4   Dim.5   Dim.6
Variance               3.074   1.638   0.687   0.329   0.237   0.035
% of var.             51.229  27.306  11.444   5.478   3.956   0.588
Cumulative % of var.  51.229  78.534  89.978  95.456  99.412 100.000

Individuals (the 10 first)
          Dist    Dim.1    ctr   cos2    Dim.2    ctr   cos2    Dim.3    ctr
LJW   |  3.502 |  1.688  0.997  0.232 | -0.419  0.115  0.014 |  0.306  0.147
LJW.1 |  1.452 |  1.128  0.445  0.604 |  0.594  0.232  0.168 | -0.644  0.649
LJW.2 |  2.521 |  0.866  0.262  0.118 |  1.892  2.350  0.563 | -1.389  3.020
LJW.3 |  0.560 |  0.044  0.001  0.006 | -0.390  0.100  0.485 | -0.313  0.153
LJW.4 |  3.148 |  1.069  0.400  0.115 |  2.767  5.026  0.773 | -0.431  0.290
LJW.5 |  1.545 | -1.026  0.368  0.441 |  0.339  0.076  0.048 | -1.094  1.876
LJW.6 |  0.899 |  0.196  0.013  0.048 | -0.082  0.004  0.008 | -0.832  1.084
LJW.7 |  1.453 | -0.219  0.017  0.023 | -0.853  0.478  0.345 |  0.961  1.445
LJW.8 |  0.819 |  0.167  0.010  0.041 | -0.546  0.196  0.445 |  0.524  0.430
LJW.9 |  6.112 |  1.021  0.365  0.028 |  4.521 13.412  0.547 |  3.411 18.220
        cos2  
LJW    0.008 |
LJW.1  0.197 |
LJW.2  0.303 |
LJW.3  0.313 |
LJW.4  0.019 |
LJW.5  0.502 |
LJW.6  0.857 |
LJW.7  0.437 |
LJW.8  0.410 |
LJW.9  0.311 |

Variables
         Dim.1    ctr   cos2    Dim.2    ctr   cos2    Dim.3    ctr   cos2  
P2O5  |  0.646 13.594  0.418 |  0.600 21.968  0.360 |  0.160  3.721  0.026 |
TiO2  |  0.921 27.594  0.848 | -0.316  6.092  0.100 |  0.052  0.399  0.003 |
K2O   |  0.889 25.723  0.791 | -0.371  8.410  0.138 |  0.114  1.903  0.013 |
Fe2O3 |  0.886 25.565  0.786 | -0.180  1.986  0.033 | -0.015  0.035  0.000 |
Cu    |  0.449  6.564  0.202 |  0.555 18.818  0.308 | -0.690 69.388  0.476 |
S     |  0.172  0.960  0.029 |  0.837 42.727  0.700 |  0.411 24.554  0.169 |
\end{verbatim}

\includegraphics{geochemistry_files/figure-pdf/PCA-4.pdf}

\begin{figure}

\centering{

\includegraphics{geochemistry_files/figure-pdf/fig-pca-1.pdf}

}

\caption{\label{fig-pca}Several pcas}

\end{figure}%

\section{Methodology}\label{methodology}

For this experiment, a Bruker portable XRF Titan S1 was used in a
laboratory benchtop setup, using battery power (up to 25\% battery
charge then replaced by a fully charged battery). A validation run was
applied on two standard samples provided by Bruker and using the
standard calibration. Samples were scanned for 180 seconds each (90
seconds for the first phase for major elements and 90 seconds for the
second phase for minor elements), at least once, with several scans
applied to samples that showed macroscopic variability (e.g., areas with
different colours or translucency). The standard database from Bruker
was used, with the Geochem application and Dual Mining method. ** chert
samples were scanned, from several sources and chert types, both
geological and archaeological. After scanning, the scanned face was
measured (thickness and diameter), to guarantee a minimum thickness and
size was followed for each sample, since other studies have shown
thickness and size may impact the homogeneity of data collection and
results (Newlander et al. 2015). All samples, including their thickness
and diameter, can be found in table X. For geological samples, fresh,
flat surfaces were scanned, avoiding altered faces or cortex; whenever
necessary, the samples were prepared by breaking the nodules. The
samples were chosen to represent all varieties of chert identified in
the Algarve region, in the archaeological record of Vale Boi, but also
from other regions such as Central Portugal and South Spain, to allow
their comparison and test hypotheses made from macroscopic and
petrographic data. For archaeological samples, artefacts were chosen
from previously identified types (REF?), focusing on larger and flatter
morphologies, with the least degree of surface alterations possible.

The analysis and result reporting protocol was established following
previous studies, focusing on the accuracy of obtained data but also the
transparency and reproducibility of the results (Newlander et al. 2015;
Johnson et al. 2024). For further reproducibility and replicability, and
working towards the goal of open science (Johnson et al. 2024; Marwick
2017), the obtained raw pXRF results can be found online on our online
compendium (LINK).

\section{Results}\label{results-1}

The pXRF measured several major and minor elements, of which a small
amount returned values of 0 or were below the limit of detection
(\textless LOD). These were uranium (U), thorium (Th), bismuth (Bi),
thallium (Tl), mercury (Hg), platinum (Pt), tantalum (Ta), hafnium (Hf),
lanthanum (La), antimony (Sb), Tin (Sn), cadmium (Cd), rhodium (Rh),
niobium (Nb), selenium (Se), zinc (Zn), nickel (Ni), cobalt (Co) and
chromium (Cr). They were removed from the analysis based on their
nonexistence, although they can still be found in the raw pXRF results.

\bookmarksetup{startatroot}

\chapter{From Stone to Tool}\label{from-stone-to-tool}

\bookmarksetup{startatroot}

\chapter{Discussion and final
remarks}\label{discussion-and-final-remarks}

\bookmarksetup{startatroot}

\chapter*{References}\label{references}
\addcontentsline{toc}{chapter}{References}

\markboth{References}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-ambroseSocialEcologicalModels1990}
Ambrose, S, and K Lorenz. 1990. {``Social and Ecological Models for the
Middle Stone Age in Southern {Africa}.''} In \emph{The Emergence of
Modern Humans: An Archaeological Perspective}, 3--33. Edinburgh:
Edinburgh University Press.

\bibitem[\citeproctext]{ref-aubryUpperPalaeolithicLithic2016}
Aubry, Thierry, Cristina Gameiro, Javier Mangado Llach, Luís Luís,
Henrique Matias, and Tiago Do Pereiro. 2016. {``Upper {Palaeolithic}
Lithic Raw Material Sourcing in {Central} and {Northern Portugal} as an
Aid to Reconstructing Hunter-Gatherer Societies.''} \emph{Journal of
Lithic Studies} 3 (2): 7--28.
\url{https://doi.org/10.2218/jls.v3i2.1436}.

\bibitem[\citeproctext]{ref-aubryEconomyLithicRaw2009}
Aubry, Thierry, and M de Araújo Igreja. 2009. {``Economy of Lithic Raw
Material During the {Upper Paleolithic} of the {C{ô}a Valley} and the
{Sic{ó} Massif} ({Portugal}): Technological and Functional
Perspectives.''} In \emph{Proceedings of the {Workshop Functional
Studies} of {Non Flint Stone Tools}: {Methodological Improvements} and
{Archaeological Inferences}}, 26. Lisbon.

\bibitem[\citeproctext]{ref-aubryRawMaterialProcurement2004}
Aubry, Thierry, X Mangado, J Fullola, L Rosell, and J Sampaio. 2004.
{``The Raw Material Procurement at the {Upper Palaeolithic} Settlements
of the {C{ô}a Valley} ({Portugal}): New Data Concerning Modes of
Resource Exploitation in {Iberia}.''} In \emph{The {Use} of {Living
Space} in {Prehistory}: {Papers} from a Session Held at the {European
Association} of {Archaeologists Sixth Annual Meeting} in {Lisbon} 2000},
37--50. Oxford: Archeopress (BAR International Series; 1224).

\bibitem[\citeproctext]{ref-belmiroProtoSolutreanLithicTechnology2020}
Belmiro, Joana. 2020. {``Proto-{Solutrean} Lithic Technology of Western
{Iberia}: The Sites of {Vale Boi} and {Lapa} Do {Picareiro}.''} Master's
thesis.

\bibitem[\citeproctext]{ref-belmiro_gravettian-solutrean_2021}
Belmiro, Joana, Nuno Bicho, Jonathan Haws, and João Cascalheira. 2021.
{``The {Gravettian-Solutrean} Transition in Westernmost {Iberia}: {New}
Data from the Sites of {Vale Boi} and {Lapa} Do {Picareiro}.''}
\emph{Quaternary International}, The 3rd {Conference World} of
{Gravettian Hunters}, 587--588 (June): 19--40.
\url{https://doi.org/10.1016/j.quaint.2020.08.027}.

\bibitem[\citeproctext]{ref-belmiroCreatingFramesReference2023}
Belmiro, Joana, Xavier Terradas, and João Cascalheira. 2023. {``Creating
Frames of Reference for Chert Exploitation During the {Late Pleistocene}
in {Southwesternmost Iberia}.''} \emph{PLOS ONE} 18 (10): e0293223.
\url{https://doi.org/10.1371/journal.pone.0293223}.

\bibitem[\citeproctext]{ref-bicho_at_2008}
Bicho, N, and J Haws. 2008. {``At the Land's End: {Marine} Resources and
the Importance of Fluctuations in the Coastline in the Prehistoric
Hunter--Gatherer Economy of {Portugal}.''} \emph{Quaternary Science
Reviews} 27 (23-24): 2166--75.
\url{https://doi.org/10.1016/j.quascirev.2008.08.011}.

\bibitem[\citeproctext]{ref-bichoEdgeCaseVale2012}
Bicho, Nuno, J. Cascalheira, and Joao Marreiros. 2012. {``On the
({L})edge:: {The Case} of {Vale Boi Rockshelter} ({Algarve}, {Southern
Portugal}).''} In \emph{Caves in {Context}. {The Cultural Significance}
of {Caves} and {Rockshelters} in {Europe}}, 65--81. Oxford: Oxbow Books.
\url{https://doi.org/10.2307/j.ctvh1djk4.10}.

\bibitem[\citeproctext]{ref-bichoPleistoceneHuntergathererCoastal2022}
Bicho, Nuno, and Eduardo Esteves. 2022. {``Pleistocene Hunter-Gatherer
Coastal Adaptations in {Atlantic Iberia}.''} \emph{Frontiers in Earth
Science} 10 (August). \url{https://doi.org/10.3389/feart.2022.957214}.

\bibitem[\citeproctext]{ref-bicho_ecodynamics_2013}
Bicho, Nuno, Tiina Manne, João Marreiros, João Cascalheira, Telmo
Pereira, Frederico Tátá, Marina Évora, Célia Gonçalves, and Leandro
Infantini. 2013. {``The Ecodynamics of the First Modern Humans in
{Southwestern Iberia}: {The} Case of {Vale Boi}, {Portugal}.''}
\emph{Quaternary International} 318: 102--16.
\url{https://doi.org/10.1016/j.quaint.2013.06.029}.

\bibitem[\citeproctext]{ref-bichoRockshelterStudiesSouthwestern2007}
Bicho, Nuno, Mary Stiner, Delminda Moura, and Armando Lucena. 2007.
{``Rockshelter Studies in Southwestern {Iberia}: The Case of Vale Boi
({Algarve}, Southern Portugal).''} In \emph{On {Shelter}'s {Ledge}:
{Histories Theories} and {Methods} of {Rockshelter Research} /{Pr{é}s}
Du Bord d'un Abri: {Les} Histories Th{é}ories Et m{é}thodes de
Recherches}, 75--82. {BAR International Series} 1655. BAR Publishing,
Oxford.

\bibitem[\citeproctext]{ref-binfordOrganizationFormationProcesses1979}
Binford, Lewis R. 1979. {``Organization and {Formation Processes}:
{Looking} at {Curated Technologies} {\textbar} {Journal} of
{Anthropological Research}: {Vol} 35, {No} 3.''} \emph{Journal of
Anthropological Research} 35 (3).
\url{https://doi.org/10.1086/jar.35.3.3629902}.

\bibitem[\citeproctext]{ref-binford_righteous_1985}
Binford, Lewis R., and Nancy M. Stone. 1985. {``{`{Righteous Rocks}'}
and {Richard Gould}: {Some Observations} on {Misguided} {`{Debate}'}.''}
\emph{American Antiquity} 50 (1): 151--53.
\url{https://doi.org/10.2307/280641}.

\bibitem[\citeproctext]{ref-bleedOptimalDesignHunting1986}
Bleed, Peter. 1986. {``The {Optimal Design} of {Hunting Weapons}:
{Maintainability} or {Reliability}.''} \emph{American Antiquity} 51 (4):
737--47. \url{https://doi.org/10.2307/280862}.

\bibitem[\citeproctext]{ref-boudagher-fadelChapterMesozoicLarger2008}
BouDagher-Fadel, Marcelle K. 2008. {``Chapter 4 {The Mesozoic} Larger
Benthic Foraminifera: The {Jurassic}.''} In \emph{Developments in
{Palaeontology} and {Stratigraphy}}, 21:157--542. Elsevier.
\url{https://doi.org/10.1016/S0920-5446(08)00004-6}.

\bibitem[\citeproctext]{ref-bousmanHuntergathererAdaptationsEconomic1993}
Bousman, Britt. 1993. {``Hunter-Gatherer Adaptations, Economic Risk and
Tool Design.''} \emph{Lithic Technology} 18 (1/2): 59--86.

\bibitem[\citeproctext]{ref-bradtmoller_lithic_2016}
Bradtmöller, Marcel, João Marreiros, Telmo Pereira, and Nuno Bicho.
2016. {``Lithic Technological Adaptation Within the {Gravettian} of the
{Iberian Atlantic} Region: {Results} from Two Case Studies.''}
\emph{Quaternary International} 406: 3--24.
\url{https://doi.org/10.1016/j.quaint.2015.08.075}.

\bibitem[\citeproctext]{ref-brantinghamMeasuringForagerMobility2006}
Brantingham, P. Jeffrey. 2006. {``Measuring {Forager Mobility}.''}
\emph{Current Anthropology} 47 (3): 435--59.
\url{https://doi.org/10.1086/503062}.

\bibitem[\citeproctext]{ref-bustilloMacroscopicClassificationFlint2009}
Bustillo, M. A., N. Castañeda, M. Capote, S. Consuegra, C. Criado, P.
Díaz-Del-Río, T. Orozco, J. L. Pérez-Jiménez, and X. Terradas. 2009.
{``Is the Macroscopic Classification of Flint Useful? {A}
Petroarcharological Analysis and Characterization of Flint Raw Materials
from the {Iberian} Neolithic Mine of {Casa Montero}.''}
\emph{Archaeometry} 51 (2): 175--96.
\url{https://doi.org/10.1111/j.1475-4754.2008.00403.x}.

\bibitem[\citeproctext]{ref-cardosoSobrePresencaLaminas2018}
Cardoso, João Luís, Marco António Andrade, and Filipe Martins. 2018.
{``{Sobre a presen{ç}a de l{â}minas de s{í}lex Ool{í}tico (e outras
mat{é}rias-primas ex{ó}genas) no povoado Calcol{í}tico do
Outeiro~Redondo (Sesimbra, Portugal): interac{ç}{ã}o durante o
3.º mil{é}nio a.C. no Sudoeste Peninsular}.''} \emph{Estudos
Arqueol{ó}gicos de Oeiras} 24: 61.

\bibitem[\citeproctext]{ref-cascalheiraTecnologiaLiticaSolutrense2010}
Cascalheira, João. 2010. \emph{{Tecnologia L{í}tica Solutrense do Abrigo
de Vale Boi (Vila do Bispo)}}. {Cadernos da UNIARQ} 5. UNIARQ.

\bibitem[\citeproctext]{ref-cascalheiraSolutrenseEmPortugal2013}
---------. 2013. {``{O solutrense em portugal: novidades do s{é}culo
xxi}.''}

\bibitem[\citeproctext]{ref-cascalheiraTerritorialityOrganizationTechnology2019}
---------. 2019. {``Territoriality and the Organization of Technology
During the {Last Glacial Maximum} in Southwestern {Europe}.''}
\emph{PLOS ONE} 14 (12): e0225828.
\url{https://doi.org/10.1371/journal.pone.0225828}.

\bibitem[\citeproctext]{ref-cascalheiraHunterGathererEcodynamics2013}
Cascalheira, João, and Nuno Bicho. 2013. {``Hunter--Gatherer Ecodynamics
and the Impact of the {Heinrich} Event 2 in {Central} and {Southern
Portugal}.''} \emph{Quaternary International} 318 (December): 117--27.
\url{https://doi.org/10.1016/j.quaint.2013.05.039}.

\bibitem[\citeproctext]{ref-cascalheira_2017}
Cascalheira, João, Nuno Bicho, Tiina Manne, and Pedro Horta. 2017.
{``Cross-Scale Adaptive Behaviors During the {Upper Paleolithic} in
{Iberia}: {The} Example of {Vale Boi} ({Southwestern Portugal}).''}
\emph{Quaternary International}, Adaptive {Cycles} in {Archaeology}, 446
(August): 17--30. \url{https://doi.org/10.1016/j.quaint.2017.01.002}.

\bibitem[\citeproctext]{ref-cascalheiraValeBoiAlgarve2013}
Cascalheira, João, Nuno Bicho, João Marreiros, Telmo Pereira, Marina
Évora, Miguel Cortés, Juan Gibaja, et al. 2013. {``Vale {Boi}
({Algarve}, {Portugal}) and the {Solutrean} in {Southwestern Iberia}.''}
\emph{Espacio Tiempo y Forma. Serie I, Prehistoria y Arqueolog{í}a} 1
(5). \url{https://doi.org/10.5944/etfi.5.2012.5376}.

\bibitem[\citeproctext]{ref-corchonrodriguezMobiliteTerritoiresRelations2016}
Corchón Rodríguez, Soledad, Jimena Martínez, and Antonio Tarriño. 2016.
{``Mobilit{é}, Territoires Et Relations Culturelles Au d{é}but Du
{Magdal{é}nien} Moyen Cantabrique : Nouvelles Perspectives.''} In
\emph{Le Concept de Territoires Dans Le {Pal{é}olithique} Sup{é}rieur
Europ{é}en}. Vol. 3, Session C16. {BAR British Archaeological Reports
International}. Oxford: Bar Publishing.

\bibitem[\citeproctext]{ref-costaLithicArrowheadsSiliceous2022}
Costa, Mafalda, Luís Dias, Leonor Rocha, Jorge Oliveira, Pedro Barrulas,
and José Mirão. 2022. {``Lithic Arrowheads: {Siliceous} Raw Material
Sources and Technology in {Southern Portugal}.''} \emph{Geoarchaeology}
37 (3): 560--73. \url{https://doi.org/10.1002/gea.21891}.

\bibitem[\citeproctext]{ref-delluniversitaDevelopmentMultiparametricCharacterisation2019}
Delluniversità, Emanuela, Italo Maria Muntoni, Ignazio Allegretta,
Massimo Tarantini, Alessandro Monno, Patrizia Maiorano, Angela Girone,
Michele Morsilli, Roberto Terzano, and Giacomo Eramo. 2019.
{``Development of a Multiparametric Characterisation Protocol for Chert
Investigation and Application on the {Gargano Promontory} Mines.''}
\emph{Archaeological and Anthropological Sciences} 11 (11): 6037--63.
\url{https://doi.org/10.1007/s12520-019-00875-8}.

\bibitem[\citeproctext]{ref-delvigneGeoresourcesTechnoculturalExpressions2019}
Delvigne, Vincent, Paul Fernandes, Peter Bindon, Jean-Pierre Bracco,
Laurent Klaric, Audrey Lafarge, Mathieu Langlais, Michel Piboule, and
Jean-Paul Raynal. 2019. {``Geo-Resources and Techno-Cultural Expressions
in the South of the {French Massif Central} During the {Upper
Palaeolithic}: Determinism and Choices.''} \emph{Anthropologica Et
Pr{æ}historica} 128/2017: 39--55.

\bibitem[\citeproctext]{ref-diazdelrioWhatIberianCopper2020}
Díaz del Río, Pedro. 2020. {``What the {Iberian Copper Age} Can Tell Us
about Peasant Societies, and Vice Versa.''} In \emph{Archaeology and
History of Peasantries}, 41--54. Documentos de Arqueolog{í}a Medieval
14, 16. Bilbao: Universidad del Pa{í}s Vasco.

\bibitem[\citeproctext]{ref-dominguezbellaMineralesRocasSociedades2010}
Domínguez Bella, Salvador. 2010. \emph{{Minerales y rocas en las
sociedades de la Prehistoria}}. C{á}diz: Grupo de Investigaci{ó}n
HUM-440, Universidad de C{á}diz.

\bibitem[\citeproctext]{ref-dominguez-bellaEstudioMateriasPrimas2006}
Domínguez-Bella, Salvador. 2006. {``El Estudio de Las Materias Primas En
La {Prehistoria} Del {Á}mbito Gaditano.''} In \emph{Actas de {I Seminaro
Hispano-Marroqu{í}} de Especializaci{ó}n En Arqueolog{í}a}, 77--88.
Universidad de C{á}diz.

\bibitem[\citeproctext]{ref-fernandesNewEvidenceConcerning2012}
Fernandes, Paulo, Jennifer A. Musgrave, Geoff Clayton, Zélia Pereira,
José Tomás Oliveira, Robbie Goodhue, and Bruno Rodrigues. 2012. {``New
Evidence Concerning the Thermal History of {Devonian} and
{Carboniferous} Rocks in the {South Portuguese Zone}.''} \emph{Journal
of the Geological Society} 169 (6): 647--54.
\url{https://doi.org/10.1144/jgs2011-156}.

\bibitem[\citeproctext]{ref-flugelMicrofaciesCarbonateRocks2010}
Flügel, Erik. 2010. \emph{Microfacies of {Carbonate Rocks}}. Berlin,
Heidelberg: Springer. \url{https://doi.org/10.1007/978-3-642-03796-2}.

\bibitem[\citeproctext]{ref-gamblePalaeolithicSocietiesEurope1999}
Gamble, Clive. 1999. \emph{The {Palaeolithic Societies} of {Europe}}.
Cambridge University Press.

\bibitem[\citeproctext]{ref-garcia-rojasGreatStepForward2021}
García-Rojas, Maite, Eder Dominguez-Ballesteros, Alejandro Prieto, Aitor
Calvo, Aitor Sánchez, Antonio Tarriño, and Alvaro Arrizabalaga. 2021.
{``A {Great Step Forward}. {Lithic Raw Material Procurement} and
{Management} Among {Palaeolithic Hunter-Gatherers} in the {Basque
Crossroads}.''} \emph{Journal of Lithic Studies} 7 (2).
\url{https://doi.org/10.2218/jls.5434}.

\bibitem[\citeproctext]{ref-gibajabaoProvenienceTechnologyMorphology2013}
Gibaja Bao, Juan Francisco, and Nuno Ferreira Bicho. 2013.
{``Provenience, Technology, Morphology and the Use of Proto-Solutrean
and Solutrean Points from {Vale Boi} ({Algarve}, Southern Portugal) :
Preliminary Results / {Provenance}, Technologie, Morphologie Et
Utilisation Des Pointes Du Proto-Solutr{é}en Et Du Solutr{é}en Du Site
de {Vale Boi} ({Algarve}, {Portugal} m{é}ridional) : R{é}sultats
Pr{é}liminaires.''}

\bibitem[\citeproctext]{ref-gomezdesolerPanadellaChertMontmaneu2020}
Gómez de Soler, Bruno Gómez, María Soto, Josep Vallverdú, Amèlia
Bargalló, M. Gema Chacón, Francesca Romagnoli, and Manuel Vaquero. 2020.
{``The {Panadella} Chert ({Montmaneu Formation}): A High-Quality Raw
Material in the {Abric Roman{í}} Sequence ({NE Iberian Peninsula}).''}
\emph{Archaeological and Anthropological Sciences} 12 (11): 252.
\url{https://doi.org/10.1007/s12520-020-01198-9}.

\bibitem[\citeproctext]{ref-gomezdesolerMultitechniqueApproachCharacterization2023}
Gómez De Soler, Bruno, María Soto, Ángel Carrancho, Francesc
Gispert-Guirado, Hans Mommsen, Juan Ignacio Morales, Alicia Muñoz Del
Pozo, et al. 2023. {``A Multi-Technique Approach to Characterization:
The {Sant Mart{í}} de {Tous} Chert as a Prehistoric Resource for the
{NE} of the {Iberian Peninsula}.''} \emph{Archaeological and
Anthropological Sciences} 15 (6): 85.
\url{https://doi.org/10.1007/s12520-023-01780-x}.

\bibitem[\citeproctext]{ref-gonzalez-samperiz_steppes_2010-1}
González-Sampériz, Penélope, Suzanne A. G. Leroy, José S. Carrión,
Santiago Fernández, Mercedes García-Antón, María José Gil-García, Paloma
Uzquiano, Blas Valero-Garcés, and Isabel Figueiral. 2010. {``Steppes,
Savannahs, Forests and Phytodiversity Reservoirs During the
{Pleistocene} in the {Iberian Peninsula}.''} \emph{Review of
Palaeobotany and Palynology} 162 (3): 427--57.
\url{https://doi.org/10.1016/j.revpalbo.2010.03.009}.

\bibitem[\citeproctext]{ref-gouldEmpiricistStrikesBack1985}
Gould, Richard A. 1985. {``The {Empiricist Strikes Back}: {Reply} to
{Binford}.''} \emph{American Antiquity} 50 (3): 638--44.
\url{https://doi.org/10.2307/280326}.

\bibitem[\citeproctext]{ref-gouldLithicProcurementCentral1985}
Gould, Richard A., and Sherry Saggers. 1985. {``Lithic {Procurement} in
{Central Australia}: {A Closer Look} at {Binford}'s {Idea} of
{Embeddedness} in {Archaeology}.''} \emph{American Antiquity} 50 (1):
117--36. \url{https://doi.org/10.2307/280637}.

\bibitem[\citeproctext]{ref-groveMovingFarMoving2023a}
Grove, Matt, Harry Hall, Lucy Timbrell, Adam Benton, and Jennifer C.
French. 2023. {``Moving Far or Moving Often? {A} Neglected Axis of
Variation in Hunter-Gatherer Mobility.''} \emph{Journal of
Archaeological Science: Reports} 52 (December): 104266.
\url{https://doi.org/10.1016/j.jasrep.2023.104266}.

\bibitem[\citeproctext]{ref-hermidaSilexNWPeninsula2016}
Hermida, Arturo De Lombera, Carlos Rodríguez Rellán, and Manuel Vaquero
Rodríguez. 2016. {``{El s{í}lex en el NW de la Pen{í}nsula Ib{é}rica. Un
estado de la cuesti{ó}n}.''} \emph{Cuadernos De Prehistoria Y
Arqueolog{í}a De La Universidad De Granada} 26: 137--55.
\url{https://doi.org/10.30827/cpag.v26i0.7398}.

\bibitem[\citeproctext]{ref-herrero-alonsoManagementLithicRaw2020}
Herrero-Alonso, Diego, Natividad Fuertes-Prieto, and Ana Neira-Campos.
2020. {``Management of Lithic Raw Materials in the {`{Mesolithic} with
Geometrics'} ({Northern} of {Iberian Peninsula}): Cha{î}nes
Op{é}ratoires and Territory.''} \emph{Journal of Archaeological Science:
Reports} 29 (February): 102093.
\url{https://doi.org/10.1016/j.jasrep.2019.102093}.

\bibitem[\citeproctext]{ref-horta_role_2019}
Horta, Pedro, João Cascalheira, and Nuno Bicho. 2019. {``The {Role} of
{Lithic Bipolar Technology} in {Western Iberia}'s {Upper Paleolithic}:
The {Case} of {Vale Boi} ({Southern Portugal}).''} \emph{Journal of
Paleolithic Archaeology} 2 (2): 134--59.
\url{https://doi.org/10.1007/s41982-019-0022-5}.

\bibitem[\citeproctext]{ref-johnsonBestPracticesPublishing2024}
Johnson, Kimberly, Colin P. Quinn, Nathan Goodale, and Richard Conrey.
2024. {``Best {Practices} for {Publishing pXRF Analyses}.''}
\emph{Advances in Archaeological Practice} 12 (2): 156--62.
\url{https://doi.org/10.1017/aap.2024.6}.

\bibitem[\citeproctext]{ref-jordaoFlintSourcesMobility2022}
Jordão, Patrícia, and Nuno Pimentel. 2022. {``Flint Sources and Mobility
at the {Chalcolithic} (3500--2200 {BCE}) Settlement of {Zambujal}
({Portugal}).''} \emph{Geoarchaeology} 37 (3): 522--43.
\url{https://doi.org/10.1002/gea.21885}.

\bibitem[\citeproctext]{ref-knuth84}
Knuth, Donald E. 1984. {``Literate Programming.''} \emph{Comput. J.} 27
(2): 97--111. \url{https://doi.org/10.1093/comjnl/27.2.97}.

\bibitem[\citeproctext]{ref-kuhnUnpackingReductionLithic1991}
Kuhn, Steven. 1991. {``{`{Unpacking}'} Reduction: {Lithic} Raw Material
Economy in the Mousterian of West-Central {Italy}.''} \emph{Journal of
Anthropological Archaeology} 10 (1): 76--106.
\url{https://doi.org/10.1016/0278-4165(91)90022-P}.

\bibitem[\citeproctext]{ref-kuhnUpperPaleolithicRaw2004}
---------. 2004. {``Upper {Paleolithic} Raw Material Economies at
{{Ü}{ç}a{ğ}{ı}zl{ı}} Cave, {Turkey}.''} \emph{Journal of Anthropological
Archaeology} 23 (4): 431--48.
\url{https://doi.org/10.1016/j.jaa.2004.09.001}.

\bibitem[\citeproctext]{ref-luedtkeArchaeologistGuideChert1992}
Luedtke, Barbara E. 1992. \emph{An {Archaeologist}'s {Guide} to {Chert}
and {Flint}}. Archaeological {Research Tools} 7. Los Angeles: University
of California.

\bibitem[\citeproctext]{ref-mannePryingNewMeaning2011}
Manne, Tiina, and Nuno Bicho. 2011. {``Prying {New Meaning} from {Limpet
Harvesting} at {Vale Boi During} the {Upper Paleolithic}.''} In
\emph{Trekking the {Shore}: {Changing Coastlines} and the {Antiquity} of
{Coastal Settlement}}. Interdisciplinary {Contributions} to
{Archaeology}. New York, NY: Springer New York.
\url{https://doi.org/10.1007/978-1-4419-8219-3}.

\bibitem[\citeproctext]{ref-manneIntensiveSubsistencePractices2012}
Manne, Tiina, João Cascalheira, Marina Évora, João Marreiros, and Nuno
Bicho. 2012. {``Intensive Subsistence Practices at {Vale Boi}, an {Upper
Paleolithic} Site in Southwestern {Portugal}.''} \emph{Quaternary
International}, Common {Problems}, {Uncommon Solutions}:
{Zooarchaeological Contributions} to {Understanding Dietary Change} in
{Mediterranean-Type Environments}, 264 (June): 83--99.
\url{https://doi.org/10.1016/j.quaint.2012.02.026}.

\bibitem[\citeproctext]{ref-manuppellaNoticiaExplicativaFolha1987}
Manuppella, G., M. Ramalho, A. Antunes, and J. Pais. 1987.
\emph{Not{í}cia Explicativa Da Folha 53-{B Tavira}}. Lisboa: Servi{ç}os
Geol{ó}gicos de Portugal.

\bibitem[\citeproctext]{ref-manuppellaNoticiaExplicativaFolha2007}
Manuppella, G., M. Ramalho, M. Antunes, and J. Pais. 2007.
\emph{{Not{í}cia explicativa da Folha 53-A Faro}}. Lisboa: Instituto
Nacional de Engenharia, Tecnologia e Inova{ç}{ã}o.

\bibitem[\citeproctext]{ref-marreirosPrimeirasComunidadesHomem2009}
Marreiros, João. 2009. {``{As primeiras comunidades do homem moderno no
Algarve Ocidental: caracteriza{ç}{ã}o paleotecnol{ó}gica e
paleoetnogr{á}fica das comunidades gravetenses e proto-solutrenses de
Vale Boi (Algarve, Portugal)}.''} Master's thesis.

\bibitem[\citeproctext]{ref-marreirosLithicTechnologyVariability2013}
Marreiros, João, and Nuno Bicho. 2013. {``Lithic Technology Variability
and Human Ecodynamics During the {Early Gravettian} of {Southern Iberian
Peninsula}.''} \emph{Quaternary International}, Paleolithic
{Ecodynamics} in southern {Iberia}, 318 (December): 90--101.
\url{https://doi.org/10.1016/j.quaint.2013.05.008}.

\bibitem[\citeproctext]{ref-marreirosLithicTechnologyGravettian2015}
Marreiros, João, Nuno Bicho, Juan Gibaja, Telmo Pereira, and João
Cascalheira. 2015. {``Lithic Technology from the {Gravettian} of {Vale
Boi}: New Insights into {Early Upper Paleolithic} Human Behavior in
{Southern Iberian Peninsula}.''} \emph{Quaternary International}, World
of {Gravettian Hunters}, 359--360 (March): 479--98.
\url{https://doi.org/10.1016/j.quaint.2014.06.074}.

\bibitem[\citeproctext]{ref-marreirosExploringLithicVariability2016}
Marreiros, João, J. Soler, Juan Gibaja, David Ortega, and N. Soler.
2016. {``Exploring Lithic Variability During the {Gravettian} in
{Iberia}: {Lithic} Technology, Use-Wear Analysis and Raw Material
Sourcing from the {Gravettian} Occupation of {Arbreda Cave}
({Catalunya}, {Spain}).''} Poster. Alcal{á} de Henares, Madrid.

\bibitem[\citeproctext]{ref-marwickComputationalReproducibilityArchaeological2017}
Marwick, Ben. 2017. {``Computational {Reproducibility} in
{Archaeological Research}: {Basic Principles} and a {Case Study} of
{Their Implementation}.''} \emph{Journal of Archaeological Method and
Theory} 24 (2): 424--50.
\url{https://doi.org/10.1007/s10816-015-9272-9}.

\bibitem[\citeproctext]{ref-marwickPackagingDataAnalytical2018}
Marwick, Ben, Carl Boettiger, and Lincoln Mullen. 2018. {``Packaging
{Data Analytical Work Reproducibly Using R} (and {Friends}).''}
\emph{The American Statistician} 72 (1): 80--88.
\url{https://doi.org/10.1080/00031305.2017.1375986}.

\bibitem[\citeproctext]{ref-matiasAprovisionamentoMateriasprimasLiticas2012}
Matias, Henrique. 2012. {``{O aprovisionamento de mat{é}rias-primas
l{í}ticas na gruta de Oliveira (Torres Novas)}.''} Master's thesis.

\bibitem[\citeproctext]{ref-matiasRawMaterialSourcing2016}
---------. 2016. {``Raw Material Sourcing in the {Middle Paleolithic}
Site of {Gruta} Da {Oliveira} ({Central Limestone Massif},
{Estremadura}, {Portugal}).''} \emph{Journal of Lithic Studies} 3 (2):
541--60. \url{https://doi.org/10.2218/jls.v3i2.1452}.

\bibitem[\citeproctext]{ref-mccallBehavioralEcologicalModels2007}
McCall, Grant S. 2007. {``Behavioral Ecological Models of Lithic
Technological Change During the Later {Middle Stone Age} of {South
Africa}.''} \emph{Journal of Archaeological Science} 34 (10): 1738--51.
\url{https://doi.org/10.1016/j.jas.2006.12.015}.

\bibitem[\citeproctext]{ref-newlanderEmpiricalStudyEffect2015}
Newlander, Khori, Nathan Goodale, George T. Jones, and David G. Bailey.
2015. {``Empirical Study of the Effect of Count Time on the Precision
and Accuracy of {pXRF} Data.''} \emph{Journal of Archaeological Science:
Reports} 3 (September): 534--48.
\url{https://doi.org/10.1016/j.jasrep.2015.07.007}.

\bibitem[\citeproctext]{ref-noceteRelacionesCentroPeriferia2005}
Nocete, Francisco, Reinaldo Sáez, José M Nieto, Rosario Cruz-Auñón,
Rosario Cabrero, Esther Álex, and Moisés R Bayona. 2005. {``{Las
relaciones centro/periferia en el Valle del Guadalquivir del III milenio
ANE. La circulaci{ó}n de hojas de caliza ool{í}tica silicificada}.''}
\emph{Tabona. Revista de Prehistoria y de Arqueolog{í}a} 14: 33--62.

\bibitem[\citeproctext]{ref-oestmoFormalModelingApproach2017}
Oestmo, Simen. 2017. {``A {Formal Modeling Approach} to {Understanding
Stone Tool Raw Material Selection} in the {African Middle Stone Age}: {A
Case Study} from {Pinnacle Point}, {South Africa}.''} PhD thesis,
ARIZONA STATE UNIVERSITY.

\bibitem[\citeproctext]{ref-oliveiraCartaGeologicaPortugal1984}
Oliveira, José Tomás. 1984. \emph{Carta {Geol{ó}gica} de {Portugal} {à}
Escala de 1/200 000. {Not{í}cia Explicativa} Da {Folha} 7.} Lisboa:
Servi{ç}os Geol{ó}gicos de Portugal.

\bibitem[\citeproctext]{ref-onezimeNewGeodynamicInterpretation2003}
Onézime, Jérôme, Jacques Charvet, Michel Faure, Jean-Louis Bourdier, and
Alain Chauvet. 2003. {``A New Geodynamic Interpretation for the {South
Portuguese Zone} ({SW Iberia}) and the {Iberian Pyrite Belt} Genesis.''}
\emph{Tectonics} 22 (August).
\url{https://doi.org/10.1029/2002TC001387}.

\bibitem[\citeproctext]{ref-ortegaMobilitatDesplacamentsDels2003}
Ortega, David. 2003. {``Mobilitat i Despla{ç}aments Dels Grups
Ca{ç}adors-Recol.lectors a Inicis Del {Paleol{í}tic Superior} a La
Regi{ó} Pirinenca Oriental.''} \emph{Cypsela: Revista de Prehist{ò}ria i
Protohist{ò}ria}, no. 14: 11--26.

\bibitem[\citeproctext]{ref-pereira_flint_2022}
Pereira, Telmo, Ana Abrunhosa, Milena Carvalho, Jonathan Haws, and
Michael Benedetti. 2022. \emph{Flint Procurement in the Western Coast of
{Iberia}: Inferring Neanderthal Mobility at {Gruta Nova} Da
{Columbeira}}.

\bibitem[\citeproctext]{ref-pereiraLithicEconomyTerritory2016}
Pereira, Telmo, Catarina Andrade, Mafalda Costa, Anne Farias, José
Mirão, and António Faustino Carvalho. 2016. {``Lithic Economy and
Territory of {Epipaleolithic} Hunter--Gatherers in the {Middle Tagus}:
{The} Case of {Pena} d'{{Á}gua} ({Portugal}).''} \emph{Quaternary
International} 412 (August): 135--44.
\url{https://doi.org/10.1016/j.quaint.2015.08.081}.

\bibitem[\citeproctext]{ref-pereira_territory_2016-1}
Pereira, Telmo, Nuno Bicho, João Cascalheira, Leandro Infantini, João
Marreiros, Eduardo Paixão, and Xavier Terradas. 2016. {``Territory and
Abiotic Resources Between 33 and 15.6 Ka at {Vale Boi} ({SW
Portugal}).''} \emph{Quaternary International} 412 (August): 124--34.
\url{https://doi.org/10.1016/j.quaint.2015.08.071}.

\bibitem[\citeproctext]{ref-pereiraRawMaterialProcurement2021}
Pereira, Telmo, Eduardo Paixão, Marina Evora, Joao Marreiros, David
Nora, Patricia Monteiro, Sandra Assis, Vânia Carvalho, and Trenton
Holliday. 2021. {``Raw Material Procurement at {Abrigo} Do {Po{ç}o Rock}
Shelter ({Central Portugal}).''} In \emph{Studies on the {Palaeolithic}
of {Western Eurasia}}. Archaeopress.

\bibitem[\citeproctext]{ref-pop_simulating_2015}
Pop, Cornel M. 2015. {``Simulating {Lithic Raw Material Variability} in
{Archaeological Contexts}: {A Re-evaluation} and {Revision} of
{Brantingham}'s {Neutral Model}.''} \emph{Journal of Archaeological
Method and Theory} 23 (4).

\bibitem[\citeproctext]{ref-ramacciottiMovingLandFirst2022}
Ramacciotti, Mirco, Oreto García-Puchol, Alfredo Cortell-Nicolau, Gianni
Gallello, Angel Morales-Rubio, and Agustín Pastor. 2022. {``Moving to
the Land: {First} Archaeometric Study of Chert Procurement at
{\emph{Cueva}}{ \emph{de La} }{\emph{Cocina}} ({Eastern Iberia}).''}
\emph{Geoarchaeology} 37 (3): 544--59.
\url{https://doi.org/10.1002/gea.21903}.

\bibitem[\citeproctext]{ref-ribeiro_evolucao_2005}
Ribeiro, Carlos. 2005. {``Evolu{ç}{ã}o {Diagen{é}tica} e
{Tectono-Sedimentar} Do {Carixiano} Da {Regi{ã}o} de {Sagres}, {Bacia
Algarvia}.''} Doctoral Thesis, {É}vora: Universidade de {É}vora.

\bibitem[\citeproctext]{ref-rochaNoticiaExplicativaFolha1983}
Rocha, R., M. Ramalho, M. Antunes, and A. Coelho. 1983. \emph{Not{í}cia
Explicativa Da {Folha} 52-{A Portim{ã}o}}. Servi{ç}os Geol{ó}gicos de
Portugal.

\bibitem[\citeproctext]{ref-rochaNoticiaExplicativaFolha1979}
Rocha, R., M. Ramalho, G. Manuppela, G. Zbyszewski, and A. Coelho. 1979.
\emph{{Not{í}cia explicativa da Folha 51-B Vila do Bispo}}. Lisbon:
Servi{ç}os Geol{ó}gicos de Portugal.

\bibitem[\citeproctext]{ref-rodriguezPrehistoricFlintExploitations2011}
Rodríguez, Antonio Morgado, José A. Losano Rodríguez, and Jacques
Pelegrin. 2011. {``{The Prehistoric Flint Exploitations of the Milanos
Formation (Granada, Spain)}.''} \emph{Journal of Andalusian Prehistory},
{MENGA 02}, 261--69.

\bibitem[\citeproctext]{ref-sanchez_goni_millennial-scale_2010}
Sanchez Goñi, Maria, and Sandy Harrison. 2010. {``Millennial-Scale
Climate Variability and Vegetation Changes During the {Last Glacial}:
{Concepts} and Terminology.''} \emph{Quaternary Science Reviews} 29
(October): 2823--27.
\url{https://doi.org/10.1016/j.quascirev.2009.11.014}.

\bibitem[\citeproctext]{ref-sanchezdelatorreNewDataChert2023}
Sánchez de la Torre, Marta, Xavier Mangado Llach, Samuel
Castillo-Jiménez, Luis Luque, José J. Alcolea-González, and Manuel
Alcaraz-Castaño. 2023. {``New Data on Chert Catchment Analysis in Inland
{Iberia} During the {Late Pleistocene}.''} \emph{Geoarchaeology}, April,
gea.21963. \url{https://doi.org/10.1002/gea.21963}.

\bibitem[\citeproctext]{ref-schmidtRapidClimateChange2012}
Schmidt, Isabell, Marcel Bradtmöller, Martin Kehl, Andreas Pastoors,
Yvonne Tafelmaier, Bernhard Weninger, and Gerd-Christian Weniger. 2012.
{``Rapid Climate Change and Variability of Settlement Patterns in
{Iberia} During the {Late Pleistocene}.''} \emph{Quaternary
International} 274 (October): 1--26.
\url{https://doi.org/10.1016/j.quaint.2012.01.018}.

\bibitem[\citeproctext]{ref-sotoProcurementMobilityLate2016}
Soto, María. 2016. {``Procurement and Mobility During the {Late
Pleistocene}: {Characterising} the Stone-Tool Assemblage of the
{Picamoixons} Site ({Tarragona}, {NE Iberian Peninsula}).''}
\emph{Journal of Lithic Studies} 3 (2): 699--724.
\url{https://doi.org/10.2218/jls.v3i2.1782}.

\bibitem[\citeproctext]{ref-surovellBehavioralEcologyLithic2009}
Surovell, Todd. 2009. \emph{Toward a {Behavioral Ecology} of {Lithic
Technology}. {Cases} from {Paleoindian Archaeology}.} Tucson: Arizona
University Press.

\bibitem[\citeproctext]{ref-tarrinoFlintRawMaterial2015}
Tarriño, Antonio, I. Elorrieta, and M. García-Rojas. 2015. {``Flint as
Raw Material in Prehistoric Times: {Cantabrian Mountain} and {Western
Pyrenees} Data.''} \emph{Quaternary International} 364 (April): 94--108.
\url{https://doi.org/10.1016/j.quaint.2014.10.061}.

\bibitem[\citeproctext]{ref-tarrinoMateriasPrimasLiticas2013}
Tarriño, Antonio, and Xavier Terradas. 2013. {``Materias Primas
l{í}ticas.''} In \emph{M{é}todos y {T{é}cnicas} de an{á}lisis y Estudio
En Arqueolog{í}a Prehist{ó}rica. {De} Lo t{é}cnico a La Reconstrucci{ó}n
de Los Grupos Humanos}. Universidad del Pa{í}s Vasco.

\bibitem[\citeproctext]{ref-terrinhaBaciaAlgarveEstratigrafia2006}
Terrinha, P, R B Rocha, J Rey, M Cachão, D Moura, C Roque, L Martins, et
al. 2006. {``{A Bacia do Algarve: Estratigrafia, Paleogeografia e
Tect{ó}nica}.''} In \emph{{Geologia de Portugal, Vol. II: Geologia
Meso-cenoz{ó}ica de Portugal.}}, 29--166. Lisbon: Livraria Escolar
Editora.

\bibitem[\citeproctext]{ref-torrenceTimeBudgetingHuntergatherer1983}
Torrence, R. 1983. {``Time Budgeting and Hunter-Gatherer Technology.''}
In \emph{Hunter-{Gatherer Economy} in {Prehistory}: {A European
Perspective}}, 11--22. Cambridge: Cambridge University Press.

\bibitem[\citeproctext]{ref-verissimo_jazidas_2004}
Verissimo, Hugo. 2004. {``Jazidas Siliciosas Da Regi{ã}o de {Vila} Do
{Bispo} ({Algarve}).''} \emph{Promontoria} 2 (2): 35--47.

\bibitem[\citeproctext]{ref-whallonSocialNetworksInformation2006}
Whallon, Robert. 2006. {``Social Networks and Information:
{Non-}{`Utilitarian'} Mobility Among Hunter-Gatherers.''} \emph{Journal
of Anthropological Archaeology}, Mesolithic mobility, exchange, and
interaction, 25 (2): 259--70.
\url{https://doi.org/10.1016/j.jaa.2005.11.004}.

\bibitem[\citeproctext]{ref-wickhamPackagesOrganizeTest2015}
Wickham, Hadley. 2015. \emph{R {Packages}: {Organize}, {Test},
{Document}, and {Share Your Code}}. O'Reilly Media.

\bibitem[\citeproctext]{ref-zambujo_jazidas_1998}
Zambujo, Ana Quelhas Gertrudes. 1998. {``{Jazidas paleol{í}ticas no
concelho de Lagos (Algarve): abordagem preliminar}.''} \emph{Revista
Portuguesa de Arqueologia} 1 (2): 5--18.

\bibitem[\citeproctext]{ref-zilhaoLatePersistenceMiddle2021}
Zilhão, João. 2021. {``The Late Persistence of the {Middle Palaeolithic}
and {Neandertals} in {Iberia}: {A} Review of the Evidence for and
Against the {`{Ebro Frontier}'} Model.''} \emph{Quaternary Science
Reviews} 270 (October): 107098.
\url{https://doi.org/10.1016/j.quascirev.2021.107098}.

\end{CSLReferences}

\cleardoublepage
\phantomsection
\addcontentsline{toc}{part}{Appendices}
\appendix

\chapter{Supplementary materials 1}\label{supplementary-materials-1}

Some supplementary materials.

\end{document}