### 2025-03-27 intro.qmd

# Introduction

This research aims to deepen our understanding of the technological, cultural, and social adaptations of hunter-gatherer groups during the Upper Palaeolithic (UP) in southwestern Europe. By examining patterns of resilience and innovation in raw material procurement and use, alongside hunter-gatherer mobility, this study investigates how these behaviours correlate with environmental changes and the evolving dynamics of social territories. The focus is on the Iberian Peninsula, particularly southwestern Iberia. Ultimately, this project will contribute to a broader understanding of how past human groups lived and adapted throughout the Upper Palaeolithic.

The UP (c. 40,000–10,000 BP) represents a key moment in human history, marked by the expansion of modern humans into Europe, significant cultural shifts, and climatic upheavals that transformed hunter-gatherer lifestyles [@bar-yosef2002; @bicho2017; @gamble2004; @straus1991; @straus1995; @sanchezgoñi2010]. These factors have made the UP a central focus in prehistoric archaeology, particularly regarding human adaptations and cultural transitions.

The Iberian Peninsula has played a critical role in this research. Its cultural and geographical specificities offer valuable opportunities to study cultural continuity and disruption through time within a single site or region.

The impact of climate on human culture, resilience, and mobility is one of the key research topics of the UP in Iberia [e.g., @bicho2017; @bicho2012; @cascalheiraHunterGathererEcodynamics2013; @morin2008; @straus1991; @bradtmoller_repeated_2012-1; @schmidtRapidClimateChange2012]. Patterns of change and adaptation have been identified primarily through lithic technological analyses, leading to the definition of the major technocomplexes that characterise the UP in Iberia: Aurignacian, Gravettian, Proto-Solutrean, Solutrean, and Magdalenian. These technocomplexes exhibit distinct knapping strategies, technological patterns, and raw material use, reflecting the technological reorganisation of past human groups. Additionally, changes in territory expansion, mobility, and social networks can be observed through variations in technological choices and the presence of long-distance raw materials [@bar-yosef2002; @zilhaoPaleoliticoSuperiorEstremadura1997; @schmidtRapidClimateChange2012; @straus1991; @fullolapericotIndustriasLiticasPaleolitico1979].

The Iberian Peninsula has often been described as a refugium during the coldest periods of the Late Pleistocene [@gomezRefugiaRefugiaPatterns2007a; @gonzalez-samperiz_steppes_2010-1; @hewitt_genetic_2000; @jochimLatePleistoceneRefugia1987; @consuegraMitochondrialDNAVariation2002; @jonesMobilitySettlementResource2013; @bichoEarlyUpperPaleolithic2017], providing milder conditions than northern Europe and supporting stable human occupation throughout the UP. However, during abrupt climatic events such as Heinrich Events and the Last Glacial Maximum, the region experienced colder temperatures, reduced precipitation, and vegetation shifts. Even within Iberia, north/south differences have been identified: during abrupt climatic events, arboreal taxa such as oak were replaced by infertile heathlands in northern Iberia or semi-desert steppe vegetation in southern Iberia [@zapataPalinologicalRecordHeinrich2003; @fletcherMillennialscaleVariabilityLast2010; @fletcherOrbitalSuborbitalscaleClimate2008; @naughtonWetDryClimatic2009; @sanchezgoniEuropeanClimaticResponse2000; @sanchezgoniContrastingImpactsDansgaard2008; @naughtonPresentdayLast252007; @roucouxCombinedMarineProxy2001; @boessenkoolNorthAtlanticSeasurface2001; @turonLandseaCorrelationsLast2003; @roucouxResponseNWIberian2005]. This would have resulted in the expansion of dry, risk-prone environments. Instead of being uniformly hospitable, Iberia may have exhibited a "leopard coat" distribution of ecological niches [@schmidtRapidClimateChange2012; @jennings_southern_2011]. These niches likely supported continuous human occupation, leading to the formation of multicomponent sites and eco-cultural niches [@yravedraNotDesertedpaleoecologyHuman2016; @hawsPaleolithicSocionaturalRelationships2012; @cascalheira_2017; @cascalheiraHunterGathererEcodynamics2013; @banksHumanEcologicalNiches2008; @forteaperezCuevaMallaetesProblemas1975].

Some studies suggest a chronological correlation between climatic shifts and the emergence of major technocomplexes, implying that cultural reorganization may have been triggered by environmental changes that altered landscapes and essential resources [e.g., @bradtmoller_repeated_2012-1; @cascalheiraHunterGathererEcodynamics2013; @bicho_rapid_2017-1]. These studies also indicate shifts in territory expansion/contraction, settlement types, mobility, and intergroup contacts [@schmidtRapidClimateChange2012]. While cultural changes are not necessarily deterministic responses to climate fluctuations, it is evident that hunter-gatherers adapted to their landscapes. Cultural, technological, and social transformations provide key insights into how and why human groups adapted at different times.

The study of raw materials, alongside lithic technological analyses, is a powerful tool for identifying patterns of change and adaptation, particularly in landscape use and mobility. Raw materials found at archaeological sites reflect human decision-making and technological organisation [@bamforth_technology_1997]. These patterns provide crucial insights into how hunter-gatherers interacted with their environments and adapted to shifting conditions [@feblot-augustinsMobilityStrategiesLate1993; @kuhn_mousterian_1995; @mellars_neanderthal_1996]. Variability in raw material use at a given site may result from factors such as mobility and procurement strategies influenced by climate [@ambroseSocialEcologicalModels1990; @binford_righteous_1985; @kuhnUnpackingReductionLithic1991; @kuhnUpperPaleolithicRaw2004; @gouldLithicProcurementCentral1985; @mccallBehavioralEcologicalModels2007], settlement patterns [@kuhnUpperPaleolithicRaw2004; @surovellBehavioralEcologyLithic2009; @groveMovingFarMoving2023a], or exchange networks and social interactions [@akermanWeaponsWunanProduction2002; @wurzHowiesonsPoortBacked1999; @whallonSocialNetworksInformation2006; @torrenceProductionExchangeStone1986].

The study of raw materials from UP sites in Iberia, particularly chert—a key raw material widely used for tool production—has provided valuable insights into hunter-gatherer procurement strategies, mobility patterns, and social networks.

A clear example is the identification of long-distance contacts between regions in Iberia, suggesting complex mobility and procurement patterns throughout the UP [e.g., @hermidaSilexNWPeninsula2016; @corchonrodriguezMobiliteTerritoiresRelations2016; @ortegaMobilitatDesplacamentsDels2003; @sanchezdelatorreNewDataChert2023]. The Côa Valley in northeastern Portugal is a particularly emblematic case, as non-local chert is present throughout the UP sequence, with potential sources located 150–250 km away in both southwestern (central Portugal) and southeastern (central Spain) directions. These patterns have been interpreted as evidence of exchange networks facilitated by seasonal movements and social interactions between groups across Iberia [@aubryFarFlintInferring2022; @aubryRawMaterialProcurement2004].

Compared to other Iberian regions, where raw material studies are more common, knowledge of UP hunter-gatherer lifeways, organisation, and social networks in southwestern Iberia has primarily been derived from lithic technology and faunal analyses. Much of this research has focused on Vale Boi, the only well-preserved UP site in the region. This presents a unique opportunity to investigate hunter-gatherer adaptations in a key area of human occupation during this period.

Vale Boi, located in southwestern Iberia, is a multi-component site spanning several UP technocomplexes, from the Gravettian (c. 32 ka cal BP) to the Magdalenian [c. 15 ka cal BP; @bichoRockshelterStudiesSouthwestern2007; @bichoEdgeCaseVale2012; @bicho_ecodynamics_2013; @cascalheira_2017; @cascalheiraHunterGathererEcodynamics2013; @manneIntensiveSubsistencePractices2012; @mannePryingNewMeaning2011]. The site, considered an eco-cultural niche, provided stable climatic conditions even during colder periods, supporting continuous residential occupations that have been consistently dated [@bicho_ecodynamics_2013; @cascalheira_2017; @belmiro_gravettian-solutrean_2021]. Extensive lithic assemblages from Vale Boi have undergone technological analysis, revealing patterns of technological change and resilience [@belmiro_gravettian-solutrean_2021; @cascalheiraValeBoiAlgarve2013; @gibajabaoProvenienceTechnologyMorphology2013; @horta_role_2019; @marreirosLithicTechnologyGravettian2015]. The site has also been identified as a centre for knowledge transfer, particularly during the Last Glacial Maximum [@cascalheiraINFLUENCIAMEDITERRANICANAS2013; @cascalheiraTerritorialityOrganizationTechnology2019].

Despite extensive technological studies on Vale Boi’s lithic assemblages, macroscopic, microscopic, and geochemical analyses of chert remain limited [@verissimo_jazidas_2004; @pereira_territory_2016-1]. This gap hinders research into raw material procurement and use, mobility, and connections between southwestern Iberia and other regions.

Consequently, Vale Boi presents an exceptional opportunity to advance understanding of UP hunter-gatherer adaptations, mobility, and technological and social organisation in Iberia.

Focusing on the lithic assemblages from Vale Boi, this research aims to investigate chert procurement, use, and mobility patterns throughout the UP of southwestern Iberia, as well as their relationship with the techno-cultural dynamics of each technocomplex and climatic changes.

To achieve this, I will address the following research questions:

-   What types of chert were exploited by hunter-gatherer groups in southwestern Portugal, and how were these materials introduced to the site?

-   How frequently did these groups move across the landscape, and can long-distance movements or social contacts be identified through the presence of non-local cherts?

-   How were different chert types managed by hunter-gatherer groups, and what impact did mobility have on their technological organisation?

-   How did technological and social organisation evolve over approximately 10,000 years of the UP in southwestern Portugal, and how might climatic changes have influenced these adaptations?

To address these questions, this research employs a multi-method approach, integrating geological surveys with macroscopic, petrographic, mineralogical, and geochemical analyses, alongside technological studies of chert lithic artefacts.

The research is structured into three peer-reviewed articles (Chapters 2, 3, and 5) and one additional chapter, each addressing different aspects of chert procurement and usage. A concluding discussion integrates the findings and examines their broader implications for understanding human behaviour and adaptation over time.

-   Chapter 2, titled "Creating Frames of Reference for Chert Exploitation during the Late Pleistocene in Southwesternmost Iberia" [@belmiroCreatingFramesReference2023] offers a geological characterisation of chert sources in the Algarve region, employing macroscopic and petrographic (thin section analysis) methods. It presents a reference collection and database of local and regional chert resources, which is crucial for distinguishing between local and non-local cherts when analysing an archaeological assemblage.

-   Chapter 3, titled "Within and Beyond: Chert Procurement Patterns During the Upper Palaeolithic in Southwesternmost Iberia" [@belmiroChertProcurementPatterns2025] focuses on the analysis of chert artefacts from the UP sequence at the archaeological site of Vale Boi, using both macroscopic and petrographic (thin section analysis) methods. These artefacts are compared with geological reference collections. The chapter offers new insights into the presence of local and non-local cherts across the various UP occupations, identifies the probable sources of non-local cherts, and explores changes in the distribution of chert types over time.

-   Chapter 4, titled "Within Chert: A Multi-Technique Mineral and Geochemical Approach to the Study of Chert in Southwestern Iberia," builds on the previous chapters by applying various techniques to analyse geological and archaeological chert samples. It provides data on the mineral and chemical composition of cherts through X-Ray Diffraction (XRD), Scanning Electron Microscopy and Energy Dispersive X-Ray Spectroscopy (SEM-EDS), and portable X-Ray Fluorescence (pXRF). These methods allow for a more comprehensive characterisation of the samples and enable the use of chemical composition to refine the source attributions of archaeological cherts.

-   Chapter 5, titled "From Stone to Tool: How Raw Materials Influenced Upper Palaeolithic Technology at Vale Boi, Southwestern Iberia," combines the previously obtained data on chert types (particularly regarding the distinction between local and non-local cherts) with technological data. It explores various patterns of chert use and management.

-   Chapter 6 synthesises the results and contributions of the previous chapters and discusses the broader implications of these patterns for understanding human behaviour.

Together, these papers and chapters present the first comprehensive study of chert raw materials associated with technology from UP occupations in southwestern Iberia and contribute to the ongoing discussion on past human adaptations.

The creation of an online database freely accessible to the scientific community provides the first online reference collection for cherts in the region. This is crucial for the development of accessible and transparent scientific production, as it offers a valuable point of reference for future studies on chert across various prehistoric periods, not only in the region but also throughout the Iberian Peninsula.