### 2025-05-06 06-discussion.qmd

# Discussion {#sec-chapter6}

In [@sec-chapter1], we outlined the research questions essential to achieving the overall goals of this doctoral dissertation and project. The answers to these questions collectively provide a detailed picture of how the Upper Palaeolithic (UP) hunter-gatherers of southwestern Iberia obtained and managed their resources, moved across the landscape, and connected with other regions of Iberia, despite their peripheral location.

In this chapter, we address those research questions by integrating and synthesising all the interpretations and conclusions presented in the previous chapters.

*What types of cherts were exploited, and how were they introduced into the site?*

Throughout the UP sequence at Vale Boi, hunter-gatherers exploited various cherts. Although this research primarily focuses on chert, it is important to note that chert constituted only a portion of the lithic raw materials used at the site. Other materials—such as quartz and greywacke—were also intensively exploited, often for specific activities [@cascalheira_2017; @pereira_territory_2016-1; @horta_role_2019].

A significant amount of cherts used at Vale Boi was local, originating from chalcedony and chert formations located inland, within approximately a 10 km radius of the site, or from coastal areas within a 20 km radius. These materials were brought to the site as nodules, likely collected from primary or sub-primary deposits, and possibly even as clusters of nodules still attached to parent rock.

No chert from more distant regional sources (e.g., \>80 km to the east) was identified in the UP sequence. This suggests that, despite the broader availability of chert throughout the Algarve region, the hunter-gatherer groups at Vale Boi exploited only a limited portion of the region’s lithic resources. Based on current geological and archaeological evidence—as well as indications of coastal resource foraging throughout the UP [@mannePryingNewMeaning2011]—it appears that the foraging radius of these groups during occupations at Vale Boi was approximately 20 km [sensu Roebroeks et al., 1988, cited in @brantinghamNeutralModelStone2003]. Chert procurement was likely embedded within broader foraging strategies [@binfordWillowSmokeDogs1980].

A portion of the cherts used at Vale Boi were non-local. In most cases, these non-local chert types are consistently present across all occupation levels and areas, excluding T11, which appears only in the Solutrean occupations, and the TL cherts, which are sporadic and represented by small groups or individual artefacts of distinct macroscopic lithotypes.

At least two different transportation strategies were identified throughout the UP sequence at Vale Boi. The majority of non-local cherts appear to have been transported to the site as nodules, partially decortified cores, or larger stone volumes, and subsequently knapped on-site. This is evidenced by the consistent presence of cortex across all non-local chert types—though notably lacking the rounded, pebble-like cortex typical of beach or river sources—as well as by the presence of cores and associated knapping by-products. These findings suggest a transportation strategy focused on larger volumes of raw material, as proposed in previous studies [@clarkPoliticsPrismaticBlades1987; @gouldLithicAssemblageWestern1971; @shottTechnologicalOrganizationSettlement1986; @kellyThreeSidesBiface1988; @perlesSearchLithicStrategies1992; @kuhnFormalApproachDesign1994; @morrowBiggerBetterComments1996], and are indicative of a site provisioning approach [@kuhnUpperPaleolithicRaw2004].

A similar model of long-distance raw material transportation is documented in France, where in the Late Gravettian hunter-gatherers transported complete or minimally pre-formed blocks over distances exceeding 200 km, in preparation for short-term hunting occupations [@delvigneGeoresourcesTechnoculturalExpressions2019].

By comparison, the Gravettian occupations of the Côa Valley exhibit intermediate long-distance raw material use patterns. While some portion of the non-local resources arrived as nodules—and are found in the archaeological record as reduced cores, core preparation, and maintenance products indicating on-site knapping—a significant portion of the imported raw materials shows reduction sequences corresponding to the final phases of bladelet production and the reuse of blanks [@aubryWeWillBe2012].

This contrasts with patterns observed at other UP sites across Iberia in relation to long-distance chert procurement. For example, at the Ametzagaina site in the western Pyrenees (Basque Country), non-local raw materials from sources located approximately 50–100 km away are present in Gravettian occupations. These materials are represented mainly by retouched tools and burin spalls, with no associated cores, although core preparation and maintenance tools are present [@arrizabalagaWhereWhat2014].

More similar to these finished-product transports, a small portion of non-local cherts at Vale Boi—possibly transported as blanks or retouched tools—was also identified throughout the UP sequence. These are represented in the TL cherts and may reflect provisioning strategies aligned with individual provisioning, as proposed by Kuhn [-@kuhnUpperPaleolithicRaw2004], involving the transport of blanks and tools as part of an individual's portable toolkit.

The presence of these varied types of non-local cherts provides evidence for mobility patterns extending beyond what would be considered the local and regional raw material catchments. While many of the non-local cherts found at Vale Boi currently do not have an identified source, at least three different sources can be surmised from chert types with macroscopic, petrographic and/or chemical similarities to known Iberian sources:

-   Cherts from the Cretaceous formations of central Portugal, found near the Lisbon and Rio Maior region [@matiasAprovisionamentoMateriasprimasLiticas2012; @matiasRawMaterialSourcing2016; @jordaoFlintSourcesMobility2022], located approximately \~215-250 km north of Vale Boi (as the crow flies). The attribution is currently based on macroscopic, petrographic and chemical similarities between archaeological and geological samples.

-   Cherts from the Upper Jurassic formations of the Betic Systems in southern Spain [@rodriguezPrehistoricFlintExploitations2011], located over \~250 km east of Vale Boi (as the crow flies). The identification is based primarily on macroscopic traits and petrographic components not found in local or regional cherts.

-   Although much less represented, the presence of jasper in the TL chert group and the distinct geochemical signature of T11, as identified through pXRF analysis, suggest these artefacts may originate from the Iberian Pyritic Belt [@araujoRegioesCentralSul2013; @oliveiraTracosGeraisEvolucao1991], located north of the Algarve region. While these resources are not available in primary deposits in the Algarve, they are known to occur in secondary deposits, such as beaches or riverbeds, which could have brought them closer to the site.

In this particular case, the jasper is found without cortex or neocortex, while T11 exhibits a platy-like cortex with an exterior fully altered by post-depositional processes. These characteristics make it difficult to determine the exact method of transportation. However, given the proximity and accessibility of secondary depositional settings—such as beaches or riverbeds—where lithic materials from the Pyritic Belt may be available, it is not possible to confidently infer the type of mobility or transport strategy involved in their acquisition.

As with the regional chert sources absent from the UP sequence at Vale Boi, the limited presence of these jaspers and other cherts in the archaeological record may reflect either their unavailability or inaccessibility during that period, or deliberate choices made by the hunter-gatherers regarding material selection and use.

*How frequently did these groups move across the landscape, and can long-distance movements or social contacts be identified through the presence of non-local cherts?*

Although long-distance cherts—originating from sources at least 200 km away—were identified at the site, the mechanisms by which they arrived at Vale Boi remain unclear. These materials may have been transported directly by the groups themselves, as part of seasonal mobility patterns, acquired through embedded or direct procurement strategies, or obtained through exchange networks maintained with other hunter-gatherer groups across Iberia.

While the UP occupations at Vale Boi have been interpreted as the result of seasonal mobility, possibly linked to deer hunting activities in spring [@manneIntensiveSubsistencePractices2012], we do not argue that the presence of cherts from over 200 km away indicates a single cultural group occupying the entire southern half of Iberia. Nor does the evidence support the idea that these materials arrived at the site solely through direct or embedded lithic procurement across such vast distances.

This interpretation is particularly relevant given the distinct directions from which non-local cherts appear to originate. For instance, a round trip between Vale Boi and central Portugal—where Cretaceous cherts were likely sourced—would amount to approximately 500 km. A similar round trip eastward to the Betic Systems of southern Spain would also be around 500 km. Together, these distances highlight the improbability of regular, direct acquisition of such materials over a total range exceeding 1,000 km.

Individually, these residential moves are feasible. Several ethnographic examples document total annual residential movements between 400 and 500 km, such as those of the Mistassini Cree [@rogersChangingSettlementPatterns1963], Tasmanians [@jonesTasmanianTribes1974], and Nukak [@politisNUKAKEthnoarchaeologyAmazonian2007]. However, such mobility often involves using aids like boats or sledges. In a broader synthesis of hunter-gatherer annual mobility, Kelly [-@kellyForagingSpectrumDiversity1995] shows that only under extreme environmental conditions—such as those faced by the Baffinland Inuit [@hantzschMyLifeEskimos1977] or the Nunamiut [@binfordNunamiutEthnoarchaeology1978]—do residential movements approach the combined distances inferred for Vale Boi when considering the northern and eastern trajectories together.

However, these distances are calculated as straight-line measurements from Vale Boi to the raw material sources, without accounting for topographic or environmental factors. Long-distance movements in those directions may have been shaped by terrain constraints, a factor incorporated into formal models of hunter-gatherer mobility [@brantinghamMeasuringForagerMobility2006].

When applied to the Iberian Peninsula, terrain friction mapping demonstrates that travel costs on foot vary significantly from direct distances measured on modern maps [@diazdelrioWhatIberianCopper2020]. In this context, distance is better understood in terms of travel time rather than absolute spatial measurements [@zilhaoLatePersistenceMiddle2021]. Díaz del Río [-@diazdelrioWhatIberianCopper2020] highlights the presence of low-cost corridors—particularly across plateaus and along the interconnected river valleys of central and southern Iberia, including the Guadalquivir, Guadiana, and Tagus. This suggests that the apparent long distances from Vale Boi to chert sources in central Portugal and southern Spain may have been less arduous and time-consuming than they seem on a flat map.

This may also be reflected in other long-distance movements identified in western Europe. In France, archaeological evidence from Gravettian seasonal hunting camps (e.g., Le Blot and Le Rond-de-Saint-Arçon) shows the exclusive use of non-local raw materials, indicating their integration into large mobility circuits within broader territorial ranges [@delvigneGeoresourcesTechnoculturalExpressions2019].

In this sense, the Gravettian occupations at Vale Boi may be interpreted within a similar framework—as seasonally occupied, short-term hunting sites used by groups originating from central Portugal and/or southern Spain, who brought with them lithic resources.

However, several technological studies of UP lithic assemblages at Vale Boi have revealed key differences in knapping strategies, raw material use, and typology when compared to assemblages from central Portugal and southern Spain [@marreirosLithicTechnologyGravettian2015]. Vale Boi has often been characterised as a site reflecting resilient behaviours and cultural continuity [@manneIntensiveSubsistencePractices2012; @cascalheira_2017], but also, particularly during the Gravettian, as exhibiting elements of cultural distinction from neighbouring groups [@marreirosLithicTechnologyGravettian2015].

While some of these differences may reflect local environmental adaptations or the specific functional characteristics of the site, the distinct typological choices—often linked to style and cultural identity—strongly suggest that the Gravettian hunter-gatherers at Vale Boi constituted a separate cultural group from those in central Portugal and southern Spain. Nevertheless, this cultural distinction does not preclude the possibility of interaction. This scenario aligns with Steward’s [-@stewardPostscriptBandsTaxonomy1969] concept of broader social networks, or the "maximum band," defining them as a group with which its members (the smaller bands of groups) vaguely identify.

Regarding raw material procurement, while it is not unheard of in ethnographic examples for one group to access the territory of another—often with permission and under established social agreements [@jonesTasmanianTribes1974; @kellyForagingSpectrumDiversity1995]—distances greater than 200 km for resource acquisition are rarely documented. Ethnographic data generally show that residential mobility seldom exceeds 300 km annually [@kellyForagingSpectrumDiversity1995], and that raw materials procured during an occupation are typically sourced within a 20 km radius [@binfordConstructingFramesReference2001]. Materials acquired from beyond this range are more often the result of transport from previous locations, rather than direct procurement during a single occupation [@surovellBehavioralEcologyLithic2009].

Therefore, it seems unlikely that the presence of distant raw materials at Vale Boi resulted from direct procurement efforts. Instead, their presence more plausibly reflects patterns of long-distance transport embedded in broader seasonal movements or social exchanges.

Archaeological examples in western Europe document levels of mobility comparable to those observed at Vale Boi, with many interpretations emphasising the role of social networks connecting hunter-gatherer groups.

For instance, UP sites in the Côa Valley (inland northern Portugal) exhibit patterns of long-distance—or "exotic"—chert acquisition strikingly similar to those at Vale Boi. Spanning the Gravettian to the Magdalenian (though with poorly preserved Solutrean occupations), the Côa Valley sites reveal non-local cherts sourced from the Rio Maior region (southwest; 130–225 km) and the Spanish Northern Meseta (east; 110–205 km). These materials have been interpreted as evidence of inter-group exchanges and the existence of social networks, with contact likely occurring along territorial boundaries during periods of aggregation—a pattern well-documented among hunter-gatherer societies [@aubryFarFlintInferring2022; @aubryWeWillBe2012; @aubryUpperPalaeolithicLithic2016; @aubryAdaptationResourcesEnvironments2015].

This interpretation is further supported by other studies linking sites in southern France (e.g., the Paris Basin) and central Portugal. During the Solutrean and Last Glacial Maximum, these sites show similar raw material procurement and technological patterns over local and regional distances, suggesting a broad interaction network spanning southwestern Iberia and western France—even during periods like the Solutrean, which are poorly represented in the Côa Valley [@aubrySolutreanLaurelLeaf2003].

Additional evidence comes from northern Iberia. At the Valverde site in Cantabria, non-local cherts sourced from over 190 km away were identified in Solutrean levels. Although rare, their presence supports the idea that UP groups in Galicia participated in an extensive network of social contacts stretching across northern Iberia and into France [@hermidaSilexNWPeninsula2016].

Similarly, the Peña Capón site in the Spanish Southern Meseta shows varying percentages of non-local cherts throughout the UP sequence, with the highest quantities during the Solutrean and decreasing presence in earlier periods like the Proto-Solutrean and Gravettian. Macroscopic analysis suggests sources located over 150 km away, interpreted as evidence for contact or mobility between the Tagus and Ebro valleys [@sanchezdelatorreNewDataChert2023].

Given these parallels, the presence of long-distance raw materials at Vale Boi can plausibly be interpreted as evidence of social interaction—possibly facilitated during seasonal aggregations [@leeIntensificationSocialLife1972], at the margins of distinct territories [@brantinghamMeasuringForagerMobility2006], or within overlapping mating networks [@wobstLocationalRelationshipsPaleolithic1976], all contributing to the reinforcement of social bonds.

Ethnographic studies further underscore the importance of social networks in explaining long-distance mobility, even when residential mobility remains limited. Long-distance movement often served key adaptive purposes, enabling the formation and maintenance of broader social ties [@gouldEmpiricistStrikesBack1985]. While a minimum band could subsist in relative isolation, multiple bands commonly participated in a larger social network to improve survival prospects and preserve cultural continuity [@wobstBoundaryConditionsPaleolithic1974]. This pattern may have been especially critical in marginal regions, where social exchange—including mating partnerships, rituals, symbolic systems, and resource trade—was essential for sustaining viable communities [@wobstLocationalRelationshipsPaleolithic1976].

Ethnographic studies demonstrate that expansive social networks are often developed by hunter-gatherers as adaptive responses to risk, especially in environments characterised by irregular and unpredictable resource availability [@gouldAnthropologyHumanResidues1978]. In such contexts, social relationships serve as mechanisms to mitigate scarcity through information exchange, mutual support, and the circulation of resources across vast areas.

For example, ethnographic accounts of !Kung hunter-gatherers show that groups or individuals travelled long distances to gather for winter-gathering, or even for long-distance exchange networks. In their exchanges, goods travelled across hundreds of miles from individual to individual [@leeIntensificationSocialLife1972]. These exchanges were not limited to utilitarian needs but played a vital role in maintaining inter-group ties. Within this framework, the acquisition of exotic goods, including lithic raw materials, can be understood as a byproduct of social cohesion rather than procurement [@wobstBoundaryConditionsPaleolithic1974; @gouldEmpiricistStrikesBack1985].

By interpreting the presence of non-local resources at Vale Boi as the result of exchange and social networks, then we assume the existence of at least two maintained social networks, previously suggested by technological studies [focused mainly in the Proto-Solutrean and Solutrean technocomplexes, @belmiro_gravettian-solutrean_2021; @cascalheiraINFLUENCIAMEDITERRANICANAS].

However, the existence of long-distance mobility in hunter-gatherer groups during the UP is, in most cases, expected. Except for situations where specific technological choices may be interpreted as ethnographic and cultural boundaries between different groups, possibly indicating small territories, the very basis of a hunter-gatherer lifeway is marked by mobility and adaptation to the shifts in natural resources and the necessity to respect their renewal times [@kellyForagingSpectrumDiversity1995].

Then, by ascertaining that southwestern Iberian hunter-gatherers moved seasonally, moved far, or at least far enough to maintain contact and exchange resources with other groups in Iberia, then there is a key follow-up question: moved how frequently?

Evidence from the site suggests that residential mobility patterns evolved throughout the UP. During the Gravettian, Vale Boi presents archaeological signals of short-term occupations, indicating higher mobility. This correlates with greater frequencies of non-local cherts, particularly those in the TL group, which may reflect either broader ranging mobility or individual provisioning strategies. In contrast, from the Proto-Solutrean onward, Vale Boi transitions into a longer-term residential settlement. Correspondingly, the frequency of non-local lithics declines, with local cherts becoming dominant during the Solutrean, consistent with reduced residential mobility [@surovellBehavioralEcologyLithic2009; @groveMovingFarMoving2023a].

*How were different chert types managed by hunter-gatherer groups, and what impact did mobility have on their technological organisation?*

These shifts in mobility frequency during the UP not only influenced the quantity of non-local raw materials found at Vale Boi but also shaped the technological and typological characteristics of the lithic assemblages. When focusing specifically on the relationship between occupation duration and technological behaviour, short-term occupations—associated with higher mobility—tend to correspond with increased frequencies of retouched tools, a pattern well-documented in ethnographic and archaeological literature [@torrenceTimeBudgetingHuntergatherer1983; @surovellBehavioralEcologyLithic2009; @groveMovingFarMoving2023a]. The more frequent occurrence of burins, for instance, may reflect an adaptive strategy to mobility, as these tools could serve as cores for the production of bladelets [@bartonGraverReconsideringBurin1996].

Across all UP phases at Vale Boi, regardless of occupation length, non-local cherts—costlier in terms of acquisition time and transport effort, even when acquired through exchange—consistently show higher proportions of retouched tools and distinct tool types. This pattern supports interpretations that non-local materials were not only more carefully curated but also reflect a higher investment in production and maintenance, aligning with theoretical expectations surrounding raw material economy [@kellyThreeSidesBiface1988; @perlesSearchLithicStrategies1992; @andrefskyRawMaterialAvailabilityOrganization1994; @kuhnUpperPaleolithicRaw2004].

The types of produced tool show diverse set of tool types, although less specialised tools, such as retouched blanks, remain predominant in non-local cherts across all assemblages. This suggests that these materials were not only transported as unspecialised tools and easily reconfigurable tools (e.g., retouched blanks) as suggested in previous studies [@odellMorphologicalExpressFunction1981; @siegelFunctionalVariabilityAssemblage1984; @shottTechnologicalOrganizationSettlement1986], but as volumes that supported tool specialisation and diversification.

In addition, non-local cherts consistently exhibit higher knapping intensity, indicating an increased effort to maximise their utility once transported to the site. This behaviour conforms to both theoretical models and archaeological expectations regarding the conservation and extended use of high-cost raw materials [@rothProductionTransportBlanks1998; @kuhnUpperPaleolithicRaw2004; @surovellBehavioralEcologyLithic2009].

This indicates that despite variations in residential mobility throughout the UP, the lithic record at Vale Boi reveals differentiated strategies of raw material management, where distance from source emerges as a key determinant in shaping technological investment and tool use.

*How did technological and social organisation evolve over approximately 10,000 years of the UP in southwestern Portugal, and how might climatic changes have influenced these adaptations?*

These results indicate a complex pattern of adaptive behaviours across the \~10,000 years of (UP) occupation at Vale Boi, shaped not only by mobility and technological strategies but also by broader climatic shifts. It is clear that both settlement and mobility patterns, as well as aspects of social organisation, evolved over time.

During the Gravettian (c. 30 ka cal BP), Vale Boi was likely occupied seasonally and in short-term occupations, with evidence of high mobility and regular contact with groups in central Portugal and southern Spain. This period seems to coincide with the onset of Heinrich Event 3 (HE 3), which brought colder, drier conditions. This would be associated with longer-term occupations of ecological niches or refugia within a generally unproductive and patchy landscape regarding resources.

However, it is key to note that the impacts of HE 3 seem to have been milder than other millennial scale, abrupt climatic events. As such, milder shifts of previous climatic conditions, especially in a Mediterranean region, may have not had a deep impact on resource availability and the landscape.

After the Heinrich stadial, which once again created mild, mediterranean conditions all throughout Iberia, the onset of the Proto-Solutrean and, subsequently, the Solutrean, coincided with Heinrich Event 2 (HE 2) and the Last Glacial Maximum (LGM) [@belmiro_gravettian-solutrean_2021; @cascalheiraHunterGathererEcodynamics2013; @gonzalez-samperiz_steppes_2010-1; @schmidtRapidClimateChange2012]. It is also during this time frame [c. 26 and 20 ka cal BP, @belmiro_gravettian-solutrean_2021; @cascalheira_2017; @cascalheiraHunterGathererEcodynamics2013] that the settlement patterns and raw material proportions begin to show alterations, after what may have been a small occupational hiatus before the Proto-Solutrean represented by lower frequencies of archaeological materials.

As such, we suggest that as climatic conditions worsened, transforming the southern Iberian landscape from Mediterranean forests to desert-like steppes [@gomezRefugiaRefugiaPatterns2007; @gonzalez-samperiz_steppes_2010-1; @hewitt_genetic_2000], hunter-gatherer groups continued seasonal mobility, but at a reduced frequency. According to Surovell’s [-@surovellBehavioralEcologyLithic2009] “Mean per Capita Occupation Span” model, this reduction likely stemmed from resource scarcity and the need to avoid overexploitation of ecological niches.

A similar pattern is observed in the UP occupation of the Côa Valley, where from the Proto-Solutrean onwards, the proportions of non-local raw materials decreased with an increased reliance on regional resources in comparison to the Gravettian occupations [@aubryWeWillBe2012]. This may indicate a similar behaviour across Iberia motivated by the occupation of ecological niches.

While it would then be expected that social networks would see themselves expanded during the HE 2 and LGM, given the scarcity of resources, and thus a higher presence of long-distance raw materials on Proto-Solutrean and Solutrean occupations, the existence of these ecological niches instead possibly fostered an environment of relative stability.

While changes in organisation and behaviour are identified, there are also key elements of resilience throughout the UP at Vale Boi. These are seen especially in the types of raw materials used throughout the UP and their differential use and maintenance. These are elements of the technological organisation of the hunter-gatherers of Vale Boi that show cultural continuity despite the changing technological choices.

The continuous presence of non-local cherts also shows that one possible element of resilience that was established from the first UP occupations of Vale Boi is the reliance on social networks. As such, the social networks established during the Gravettian appear to have persisted throughout the UP sequence at Vale Boi, supporting the notion of stable intergroup ties for at least 10,000 years.

Ethnographic and archaeological parallels further support this interpretation. At Puntutjarpa Rockshelter in Australia's western desert, hunter-gatherers continuously occupied the site for over 10,000 years, relying on local lithic sources but maintaining low-frequency access to exotic raw materials (\>40 km away) throughout the sequence—clear evidence of a stable long-distance social network [@gouldAnthropologyHumanResidues1978]. In Iberia, the Côa Valley presents a similar case, with a social and resource network extending up to 400 km and maintained throughout the UP [@aubryWeWillBe2012].

In sum, the patterns of raw material use at Vale Boi—interwoven with changes in settlement duration, tool production, and environmental context—offer a nuanced understanding of UP mobility and social strategies. These findings not only refine previous interpretations of the site but also highlight the resilience and adaptability of the human groups who occupied it.